---
title: "Tidy Tuesday 2021 Week 40"
date: "2021/09/28"
output: html_notebook
---

[TidyTuesday](https://github.com/rfordatascience/tidytuesday) week 40: [NBER papers](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-09-28/readme.md), data from the National Bureau of Economic Research [NBER](https://www2.nber.org/RePEc/nbr/nberwo/) by way of the [nberwp package by Ben Davies](https://github.com/bldavies/nberwp).

```{r}
library(tidyverse)
library(glue)
library(gt)
library(gtExtras)
```

```{r}
papers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/papers.csv')
authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/authors.csv')
programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/programs.csv')
paper_authors <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_authors.csv')
paper_programs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-28/paper_programs.csv')
```

```{r}
joined_df <- left_join(papers, paper_authors) %>% 
  left_join(authors) %>% 
  left_join(paper_programs) %>% 
  left_join(programs)%>% 
  mutate(
    catalogue_group = str_sub(paper, 1, 1),
    catalogue_group = case_when(
      catalogue_group == "h" ~ "Historical",
      catalogue_group == "t" ~ "Technical",
      catalogue_group == "w" ~ "General"
    ),
    .after = paper
  ) 
```

### Working paper count by program and decade
* shared on [Twitter](https://twitter.com/leeolney3/status/1442671857015087106)
* inspired by https://bldavies.com/blog/female-representation-collaboration-nber/

```{r}
wp = joined_df %>% group_by(program, program_desc, year) %>%
  summarise(n=n_distinct(paper)) %>%
  arrange(year) %>%
  mutate(decade=case_when(between(year,1980,1989)~"1980s",
                          between(year,1990,1999)~"1990s",
                          between(year,2000,2009)~"2000s",
                          between(year,2010,2019)~"2010s"
                          )) %>%
  drop_na() 

wp2 = wp %>% mutate(program=glue::glue("{program_desc} ({program})")) %>%
  group_by(program) %>% mutate(total=sum(n)) %>%
  arrange(year, program) %>%
  mutate(spark=list(n)) %>%
  select(program, spark, total) %>%
  distinct()

wp3 = wp %>% mutate(program=glue::glue("{program_desc} ({program})")) %>%
  group_by(program, decade) %>% tally(n) %>% 
  ungroup() %>%
  pivot_wider(names_from = decade, values_from=n) %>%
  mutate_if(is.numeric, list(~replace_na(., 0)))
```


```{r}
wp3 %>% inner_join(wp2, by="program") %>%
  select(Program=program, Total=total, "1980s","1990s","2000s","2010s",Trend=spark) %>%
  arrange(desc(Total)) %>%
  gt() %>%
  gt_theme_espn() %>%
  cols_align(Program, align="left") %>%
  gt_plt_dot(Total, Program,palette = "rcartocolor::ag_GrnYl", max_value=5246) %>%
  gtExtras::gt_sparkline(Trend) %>%
  tab_options(table.font.size = 12.5,
              heading.subtitle.font.size = 14) %>%
  gt_color_box(`1980s`, domain=2:786) %>%
  gt_color_box(`1990s`, domain=2:797) %>%
  gt_color_box(`2000s`, domain=132:1647) %>%
  gt_color_box(`2010s`, domain=200:2424) %>%
  tab_header(title="Economic Papers", subtitle="Working papers count by program and decade") %>%
  tab_source_note(source_note="TidyTuesday Week 40 | Data source: National Bureau of Economic Research (NBER) by way of the nberwp package by Ben Davies") 
```

#### Count and percent of female authors by year

```{r}
library(gender)
library(hrbrthemes)
library(ggtext)
library(patchwork)
library(scales)
library(colorspace)
```

```{r}
# add gender
paper_authors<-paper_authors%>%
  left_join(papers)%>%
  left_join(authors)%>%
  separate(name,c("given",NA))
name<-gender(paper_authors$given, years = c(2020-60,2020-30))
paper_authors<-paper_authors%>%
  left_join(name,by= c("given"="name"))
```

```{r}
female = paper_authors %>% 
  filter(between(year, 1980, 2020)) %>%
  group_by(year,gender) %>%
  tally() %>%
  mutate(prop=n/sum(n)) %>%
  filter(gender=="female")
```

```{r}
# count and percent of female authors
female %>% 
  mutate(prop=round(prop*100,2)) %>%
  rename("Count"=n,"Percent"=prop) %>%
  pivot_longer(Count:"Percent") %>%
  mutate(value=ifelse(name=="Percent",-1*value,1*value)) %>%
  ggplot(aes(x=year, y=value, fill=name, alpha=ifelse((value==max(value)|value==min(value)),1,0.6))) +
  scale_alpha_identity() +
  scale_fill_manual(values=c("#31572c","#034078")) +
  geom_col() +
  facet_grid(str_to_upper(name)~., scales="free",switch="both") +
  #scale_x_continuous(expand=c(0.01,0.01)) +
  scale_y_continuous(labels=abs) +
  theme_ipsum(base_size = 8) +
  theme(panel.spacing = unit(0, "lines"),
        panel.grid.minor=element_blank(),
        strip.placement = "outside",
        strip.text.y.left=element_text(angle=0, size=8, hjust=.5),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position = "none",
        plot.title.position = "plot",
        plot.title=element_text(size=13, face="plain"),
        plot.subtitle=element_markdown(size=10, color="#5c6b73"),
        plot.caption.position="plot",
        plot.caption=element_text(face="plain",size=7, hjust=0),
        plot.margin = unit(c(1, 2, 1, 1), "cm")
        ) +
  labs(title="Female Authors in NBER Papers",
       caption="Data source: National Bureau of Economic Research (NBER) by way of the nberwp package by Ben Davies",
       subtitle= "<span style = 'color:#31572c;'><b>Count</b></span> and <span style = 'color:#034078;'><b>Percent</b></span> of female authors in NBER papers by year, from 1980 to 2020")
```

### Percentage of program by gender

```{r}
gender_df = joined_df %>% separate(name,c("given",NA))
name<-gender(gender_df$given, years = c(2020-60,2020-30))
gender_df<-gender_df %>%
  left_join(name,by= c("given"="name"))
```


```{r}
gender_tab = gender_df %>% 
  select(program_category, program_desc, gender) %>%
  drop_na() %>%
  count(program_category, program_desc, gender)
```

```{r}
gt2 = gender_tab %>%
  group_by(gender) %>%
  mutate(prop=n/sum(n)) %>%
  mutate(rank = rank(prop, ties.method = "random"))

gt2 %>%
  ggplot(aes(x=str_to_title(gender), y=rank, color=program_category)) +
  geom_point(size=5) +
  geom_line(aes(group=program_desc), alpha=.5) +
  scale_color_manual(values=c("#219ebc","#e85d04","#023047")) +
  geom_text(data=gt2 %>% filter(gender=="female"), aes(label=program_desc), size=2.5, hjust=1, nudge_x = -.07,family="Arial Narrow") +
  geom_text(data=gt2 %>% filter(gender=="male"), aes(label=program_desc), size=2.5, hjust=0, nudge_x = .07, family="Arial Narrow") +
  geom_text(data=gt2 %>% filter(gender=="female"), aes(label=round(prop*100,1)), 
            size=2.2, color="white",fontface="bold",family="Arial Narrow") +
  geom_text(data=gt2 %>% filter(gender=="male"), aes(label=round(prop*100,1)), 
            size=2.2, color="white",fontface="bold",family="Arial Narrow") +
  scale_x_discrete(position="top", expand=c(0.6,0.6)) +
  theme_minimal(base_size = 10, base_family = "Arial Narrow") +
  guides(color=guide_legend(reverse=T)) +
  theme(legend.margin=margin(t=-5),
        axis.title=element_blank(),
        panel.grid = element_blank(),
        axis.text.y=element_blank(),
        axis.text.x.top=element_text(size=10, margin=margin(b=-2), color="black"),
        legend.text=element_text(size=8),
        legend.title=element_text(size=8),
        plot.margin = unit(c(.5, .5, .5, .5), "cm"),
        plot.title=element_text(margin=margin(b=13))) +
  labs(color="Program Category",
       title="NBER Papers: Percentage of programs by gender")
```

#### Count and percent of females by program 

```{r}
df = joined_df %>% 
  select(name, program_desc, year) %>% 
  drop_na(program_desc, year) %>%
  separate(name, into = c('name',NA))

gender_names <- gender(df$name) %>% 
  select(name, gender) %>% 
  distinct()

gender_pd = df %>% 
  count(name, program_desc) %>%
  left_join(gender_names) %>% 
  na.omit() %>%
  group_by(program_desc, gender) %>% 
  summarise(n = sum(n)) %>% 
  mutate(prop = n/sum(n)) %>% 
  ungroup()
```

```{r}
gender_pd %>%
  filter(gender=="female") %>%
  ggplot(aes(y=reorder(program_desc,prop), x=prop)) +
  geom_vline(xintercept = 0, size=.2, color="#adb5bd") +
  geom_richtext(aes(x=0.001,label=glue::glue("{program_desc} | **{percent(prop, accuracy=.1)}**")),
            size=1.9, hjust=0, nudge_y = 0.35,fill=NA, label.color=NA) +
  geom_text(aes(label=n), size=1.8, hjust=-0.75) +
  geom_segment(aes(x=0, xend=prop, y=reorder(program_desc,prop), yend=reorder(program_desc,prop)), color="#05668d") +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_discrete(expand = expansion(mult = c(0, .035))) +
  geom_point(aes(size=n), show.legend = F, color="#f25c54", alpha=.8) +
  coord_cartesian(clip="off") +
  theme_void(base_size = 9) +
  theme(plot.margin = unit(c(.5, 3, .5, 3), "cm"),
        plot.subtitle = element_markdown(size=7, lineheight = 1.3),
        plot.title=element_markdown(size=9.5),
        plot.caption=element_text(size=5)) +
  labs(subtitle = "by program, bar length represents <b><span style = 'color:#05668d;'>percent</span></b> and circle size represents <b><span style = 'color:#f25c54;'>count</span></b> of female authors<br>",
       title="Female Authors in NBER Publications",
       caption="\nData source: National Bureau of Economic Research")
```

#### Proportion of females authors by program and year
* inspired by: https://twitter.com/toeb18/status/1443498109846368256

```{r}
gender_pd2 = df %>% 
  count(name, program_desc,year) %>%
  left_join(gender_names) %>% 
  na.omit() %>%
  group_by(program_desc, year, gender) %>% 
  summarise(n = sum(n)) %>% 
  mutate(prop = n/sum(n)) %>% 
  ungroup() %>% filter(gender=="female")

#gender_pd2 %>% group_by(program_desc, year) %>% mutate(paper=sum(n))

female2020 = gender_pd2 %>% 
  filter(year==2020) %>% 
  arrange(desc(prop)) %>% 
  mutate(program_desc=fct_inorder(program_desc)) 

gender_pd2$program_desc = factor(gender_pd2$program_desc, levels = female2020$program_desc)
```

```{r, fig.width=3.5, fig.height=4}
gender_pd2 %>% 
  ggplot(aes(y=fct_rev(program_desc), x=year)) +
  geom_point(aes(color = prop, size = prop), shape = 15) +
  geom_text(data = gender_pd2 %>% filter(year==2020), aes(x=1975,label=program_desc), 
            size=2.2, hjust=0, vjust=-1.7) +
  scale_x_continuous(expand=c(0.02,0.02)) +
  scale_size_continuous("% of Female Authors", range = c(1, 4), 
                        breaks=c(0.02,0.25, 0.50,0.75,1.00),labels=percent_format())+
  scale_color_continuous_sequential(palette="viridis", name="% of Female Authors",
                                    breaks=c(0.02,0.25, 0.50,0.75,1.00),labels=percent_format()) +
  guides(color= guide_legend(), size=guide_legend(), alpha = guide_legend()) +
  theme_minimal(base_size = 7) +
  theme(axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank(),
        legend.direction = "horizontal",
        axis.line.x.bottom = element_line(),
        legend.position = "bottom",
        plot.title=element_markdown(size=12),
        plot.subtitle=element_text(size=8, color="#343a40"),
        plot.margin = unit(c(.7, 1, .5, .5), "cm")
        ) +
  labs(title="Proportion of **Female** Authors in NBER Publications",
       subtitle="Publication history in different programs, from 1975 to 2020\n",
       x="Publication Year")
  
```




