---
title: "Tidy Tuesday Week 25/2021"
output: html_notebook
---

[Tidy Tuesday](https://github.com/rfordatascience/tidytuesday) week 25 [Du Bois and Juneteenth Revisited](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-06-15/readme.md), data from [Anthony Starks](https://twitter.com/ajstarks), [Allen Hillery](https://twitter.com/AlDatavizguy/status/1358454676497313792?s=20) and [Sekou Tyler](https://twitter.com/sqlsekou/status/1360281040657522689?s=20).


```{r}
# Load libraries
library(tidyverse)
library(ggtext)
library(patchwork)
library(lubridate)
library(ggsci)
library(plyr)
library(countrycode)
library(sp)
library(rworldmap)
library(ggsci)
```


```{r}
# import data 
tweets <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-15/tweets.csv')

glimpse(tweets)
```


```{r}
# clean
tweets1 = tweets %>% 
  dplyr::select(long, lat) %>% drop_na()
tweets1

# function to get country 
coords2country = function(points)
{  
  countriesSP <- getMap(resolution='coarse')
  pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))  
  indices = over(pointsSP, countriesSP)

  #indices$continent   # returns the continent (6 continent model)
  #indices$REGION   # returns the continent (7 continent model)
  indices$ADMIN  #returns country name
  #indices$ISO3 # returns the ISO3 code 
}

# get country 
tweets1$country = coords2country(tweets1)
glimpse(tweets1) 
```



```{r}
# join
tweets_joined = join(tweets, tweets1,type="left",match="first")
# get region (7 Regions as defined in the World Bank Development Indicators)
tweets_joined$region = countryname(tweets_joined$country,destination = "region", warn = FALSE)
tweets_joined$date = date(tweets_joined$datetime)
glimpse(tweets_joined)
```


```{r}
# wrangle 
plot_data = tweets_joined %>% 
  group_by(date, region) %>% 
  dplyr::summarise(unique_users = n_distinct(username),total_tweets=n()) %>% 
  mutate(region_NA = ifelse(region=="North America", "North America","All other regions")) %>%
  mutate(region_NA = ifelse(is.na(region_NA),"No location",region_NA)) %>%
  filter(!is.na(date))

summary(plot_data$unique_users)
```


```{r, fig.height=3, fig.width=3.5}
# plot
plot_data %>%
  ggplot(aes(x=date, y=total_tweets, size=unique_users, 
             color=factor(region_NA,levels=c("No location","All other regions","North America"),ordered=T))) + 
  geom_jitter(alpha=0.82, width=0.4) + 
  scale_size_continuous(breaks=c(1,8,18,28)) + 
  theme_minimal(base_size=10) +
  guides(color=FALSE) + 
  scale_color_manual(values =c("#6c77a1","white","#de2d48")) +
  theme(legend.position="top",
        panel.grid.minor=element_blank(),
        panel.grid.major=element_line(linetype = "13", color="grey65"),
        plot.title=element_text(face="bold",size=24),
        plot.subtitle=element_markdown(size=13),
        axis.title=element_text(face="bold", size=9.5),
        plot.background=element_rect(fill="#ead4bc", color=NA),
        panel.background = element_rect(fill="#ead4bc", color=NA),
        plot.margin = ggplot2::margin(20, 30, 20, 20),
        legend.title=element_text(size=9.5)) +
  labs(x="Date", y="Tweet Count", size="Unique Users", color="Region",
       caption="\nTidy Tuesday Week 25 | Data from Anthony Starks, Allen Hillery and Sekou Tyler",
       title="#DuBoisChallenge Tweets",
       subtitle="<span style = 'color:#de2d48'><b>North America</b></span>, <span style = 'color:#6c77a1'><b>All other regions</b></span> and <span style = 'color:#FFFFFF'><b>No location</b></span><br>") 
```


ALT text: Scatterplot showing #DuBoisChallenge total tweets count by date, the dot size shows the number of unique users and color indicates location (North America, all other regions and no location). The plot shows that total number of tweets and unique users from North America is consistently higher than all other regions and no location. 


```{r, warning=F}
# region and date
tweets_joined %>% group_by(date, region) %>% 
  filter(!is.na(region)) %>%
  ggplot(aes(y=region, x=date)) + 
  geom_point(aes(color=region), show.legend=F) + 
  facet_wrap(~region, ncol=1, scale="free_y") + 
  theme_minimal(base_size=10) + 
  theme(strip.text=element_blank(),
        plot.title.position = "plot") + 
  scale_color_npg() + 
  labs(x="Date", y="Region", title = "#DuBoisChallenge tweets by region and date")  
```



