---
title: "Wealth and Income"
date: "2021-02-09"
output: html_notebook
---

## Data Visualization Exercise

[__TidyTuesday__](https://github.com/rfordatascience/tidytuesday) 2021/week7 

[Wealth and Income over time](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-02-09/readme.md), data from the [Urban Institute](https://apps.urban.org/features/wealth-inequality-charts/) and the [US Census](https://www.census.gov/data/tables/time-series/demo/income-poverty/historical-income-households.html).


### Load libaries
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(ggtext)
library(gridExtra)
library(colorspace) 
library(lemon) # reposition legend 
library(gggibbous) # moon line plot
library(ggstream) #ridge stream plot
library(CGPfunctions) #slope chart
```
 
 
### Import Data
```{r}
tuesdata <- tidytuesdayR::tt_load(2021, week = 7)

lifetime_earn <- tuesdata$lifetime_earn
student_debt <- tuesdata$student_debt
retirement <- tuesdata$retirement
home_owner <- tuesdata$home_owner
race_wealth <- tuesdata$race_wealth
income_time <- tuesdata$income_time
income_limits <- tuesdata$income_limits
income_aggregate <- tuesdata$income_aggregate
income_distribution <- tuesdata$income_distribution
income_mean <- tuesdata$income_mean
```


### P1: student_debt (percentage) 
* shared on [twitter](https://twitter.com/leeolney3/status/1358864374421741570/photo/1) 

```{r}
# loan_debt_pct
student_debt %>% 
  ggplot(aes(x=year, y= loan_debt_pct, color=race)) + 
  geom_line(size=1) + 
  scale_x_continuous(limits=c(1989, 2016.9), breaks=seq(1989,2016.9,3)) + 
  scale_y_continuous(limits=c(0,0.45), breaks=seq(0,0.45,0.1), labels=percent_format(accuracy=1)) +
  labs(x="",
       y="",
       title="Share of Families with Student Loan Debt in America",
       subtitle= "The percentage of <span style = 'color:#9e2a2b'><b>Black</b></span>-families with student loan debt <br> has been higher than <span style = 'color:#fb8500'><b>White</b></span>-families and  <span style = 'color:#006d77'><b>Hispanic</b></span>-families since 2001",
       caption="Data from Urban Institute and US Census") +
  scale_color_manual(values=c("#9e2a2b","#006d77","#fb8500")) + 
  theme_light() +
  theme(panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        legend.position= "none",
        plot.title=element_text(size=14,face="bold"),
        plot.subtitle=ggtext::element_markdown()
        ) + 
  geom_text(data=student_debt,aes(y=0.43, x=2015, label="Black"), color="#9e2a2b", size=3) + 
  geom_text(data=student_debt,aes(y=0.345, x=2015, label="White"), color="#fb8500", size=3) + 
  geom_text(data=student_debt,aes(y=0.23, x=2014.8, label="Hispanic"), color="#006d77", size=3) +
  coord_cartesian(expand=FALSE)
```

### P2: student_debt (average)  
* geom_textbox and geom_richtext reference: [ggplot Wizardry Hands-On by Cédric Scherer](https://z3tt.github.io/OutlierConf2021/)

```{r}
# loan_debt 
p1a = student_debt %>% 
  ggplot(aes(x=year, y= loan_debt, color=race)) + 
  geom_line(size=1) + 
  scale_x_continuous(limits=c(1989, 2016), breaks=seq(1989,2016,3)) + 
  scale_y_continuous(limits=c(0,15000), breaks=seq(0,15000, 2500), labels=scales::dollar_format()) +
  scale_color_manual(values=c("#9e2a2b","#006d77","#fb8500")) +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(),
        legend.position="none",
        plot.title=element_text(size=14,face="bold"),
        plot.subtitle = element_text(size=10)) + 
  # add rich text
  ggtext::geom_richtext(
    data=tibble(
    x=c(2014.5,2014.5,2014.5),y=c(12900,10200,6000),
    race1= c("Black","White","Hispanic"),
    lab = c("<b style='font-size:12pt;'>Black</b>", 
            "<b style='font-size:12pt;'>White</b>", 
              "<b style='font-size:12pt;'>Hispanic</b>"),
      angle = c(50, 42, 51)
    ), 
  aes(x,y,label=lab, color=race1,angle=angle),
  size=3, fill=NA, label.color=NA, 
  lineheight=.3) +
  labs(title="Student Loan Debt In America",
       subtitle=" Average family student loan debt for aged 25-55, by race and year normalized to 2016 dollars",
       caption="Data from Urban Institute and US Census",
       y="",
       x="")
```


```{r}
p1b = p1a + 
  theme(
    ## turn title into filled textbox
    plot.title = ggtext::element_textbox_simple(
      color = "white", fill = "#1d3557",  size = 14, 
      padding = margin(8, 4, 8, 4), margin = margin(b = 5), lineheight= .9
    ),
    ## add round outline to caption
    plot.caption = ggtext::element_textbox_simple(
      width = NULL, linetype = 1, color="slategrey", padding = margin(3, 8, 3, 8), 
      margin = margin(t = 15), r = grid::unit(8, "pt")
    ),
    axis.title.x = ggtext::element_markdown(),
    axis.title.y = ggtext::element_markdown(),
  ) +
  ## textbox
  ggtext::geom_textbox(
    data = tibble(x = 1995, y = 12200, label = "From 2007 to 2016, the average student loan debt of <span style = 'color:#9e2a2b'><b>Black</b></span>-families is higher than <span style = 'color:#fb8500'><b>White</b></span>-families and  <span style = 'color:#006d77'><b>Hispanic</b></span>-families."),
    aes(x, y, label = label), inherit.aes = FALSE,
    size = 3,
    fill = "#e5e5e5", box.color = "#457b9d",
    width = unit(11, "lines")
  ) 

p1b
```

### P3: student_debt (loan_debt and loan_debt_pct)

```{r}
student_debt %>% 
  ggplot(aes(x=loan_debt, y= loan_debt_pct, color=race)) +
  geom_point(aes(shape=race),size=2) + 
  theme_minimal() + 
  scale_color_manual(values=c("#9e2a2b","#006d77","#fb8500")) + 
  scale_y_continuous(label=percent_format(accuracy=1)) +
  labs(title= "Student loan debt and share of families with student loan debt",
       color="Racial group",
       shape="Racial group",
       caption="Data from Urban Institute and US Census") +
  theme(legend.position="top",
        plot.title=element_text(face="bold",hjust=0.5),
        plot.caption=element_text(color="slategrey"))
```

### P4: race_wealth 

```{r}
rw = race_wealth %>% 
  filter(!is.na(wealth_family)) %>% 
  filter(year>=1989) %>% 
  filter(type=="Median") %>%
  filter(race!="Non-White") %>%
  mutate(wealth=round(wealth_family)) %>%
  mutate(grp = ifelse(wealth_family>100000,"Y","N"))
```

```{r}
rw %>%
  ggplot(aes(x=as.factor(year), race)) +
  geom_tile(aes(fill=wealth_family), color="white", size=2) + 
  geom_text(aes(label=comma(wealth), color=grp),size=3) +
  theme_minimal() +
  theme(legend.position="bottom",
        axis.ticks = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y=element_text(face="bold"),
        legend.title = element_text(size=10, color="black"),
        plot.title=element_text(face="bold"),
        plot.title.position = "plot",
        plot.subtitle = element_text(),
        plot.caption=element_text(hjust=0, color="slategrey"),
        plot.caption.position = "plot"
        ) + 
  scale_x_discrete(position="top") + 
  scale_fill_continuous_sequential(palette="Dark Mint", trans="log10", label=dollar) +
  scale_color_manual(values=c("black","white")) + 
  labs(x="",
       y="",
       fill="Median Family Wealth",
       title= "Family Wealth in America by Race and Year",
      subtitle = "Normalized to 2016 dollars",
      caption = "Data from Urban Institute and US Census") +
  guides(fill= guide_colorbar(title.position="top",
                              title.hjust = .5,
                              barwidth = unit(20, "lines"),
                              barheight = unit(.5, "lines"))) +
  guides(color=FALSE) #remove text color legend 
```

### P5: lifetime earn 
* plot design reference: [Tobias Stalder](https://twitter.com/toeb18/status/1358408394722394117/photo/1) 
```{r}
lifetime_earn %>% 
  ggplot(aes(x=lifetime_earn, y=fct_rev(race))) +
  geom_line(aes(group=race),size=7,color="#b8b8d1",alpha=0.5) + 
  geom_point(aes(color=gender),size=8,alpha=0.9) + 
  scale_x_continuous(labels=scales::dollar_format(), limits=c(1000000, 3000000)) +
  theme_minimal() + 
  scale_color_manual(values=c("#3a6ea5","#bd632f")) +
  theme(legend.position="none",
        axis.text.y=element_blank(),
        plot.title = element_text(face="bold",hjust=0.5)) +
  geom_text(aes(x=1200000, y=3.25, label="Female"), color="#bd632f",size=3.6) + 
  geom_text(aes(x=1800000, y=3.25, label="Male"), color="#3a6ea5",size=3.6) + 
  geom_text(aes(x=1500000, y=3, label="Black"),size=3.5, family="Courier") + 
  geom_text(aes(x=1550000, y=2, label="Hispanic any race"),size=3.5, family="Courier") +
  geom_text(aes(x=2100000, y=1, label="White"),size=3.5, family="Courier") + 
  labs(y="",
       x="",
       title="Average lifetime earning in America by race and gender",
       caption="Data from Urban Institute and US Census")
```


### P6: home_owner
```{r}
summary(home_owner$year)
```

```{r}
home_owner %>% 
  filter(year>2006) %>% 
  ggplot(aes(x=year, y=home_owner_pct)) +
  geom_line(aes(color=race), size=1.5, alpha=0.8) +
  geom_moon(data= (home_owner %>% filter(year>2006)), aes(ratio=1),size=5, fill="#adb5bd",color="white") +
  geom_moon(data= (home_owner %>% filter(year>2006)), aes(ratio=home_owner_pct, fill=race), size=5,color="white") + 
  scale_y_continuous(limits=c(0.35,0.75),labels= percent_format(accuracy=1)) +
  scale_x_continuous(breaks=seq(2007, 2016,1)) +
  theme_minimal() +
  theme(panel.grid.minor=element_blank(),
        legend.position="none",
        panel.grid=element_line(size=0.3),
        plot.caption = element_text(size=8, color="slategrey",hjust=0.5),
        plot.title = element_text(face="bold", hjust=0.5)) + 
  scale_color_manual(values=c("#9e2a2b","#006d77","#fb8500")) + 
  scale_fill_manual(values=c("#9e2a2b","#006d77","#fb8500")) + 
  geom_text(data=(home_owner %>% 
  filter(year>2006)),aes(y=0.71, x=2015, label="White"), color="#fb8500", size=3.5) + 
  geom_text(data=(home_owner %>% 
  filter(year>2006)),aes(y=0.49, x=2015, label="Hispanic"), color="#006d77", size=3.5) +
  geom_text(data=(home_owner %>% 
  filter(year>2006)),aes(y=0.39, x=2015, label="Black"), color="#9e2a2b", size=3.5) + 
  labs(x= "",
       y= "",
       title= "Home Ownership Percentage in America (2007 to 2016)",
       subtitle=" by year and racial group",
       caption = "Data from Urban Institute and US Census") + 
  theme(plot.title = ggtext::element_textbox_simple(
        color = "white", fill = "#1d3557",  size = 14, 
        padding = margin(8, 4, 8, 4), margin = margin(b = 5), lineheight= .9))
```

### P7: mean_income 

```{r}
im2 = income_mean %>%
  filter(income_quintile != "Top 5%") %>%
  mutate(income_quintile = factor(income_quintile, levels = c("Lowest", "Second", "Middle", "Fourth", "Highest"))) %>%
  filter(dollar_type == "2019 Dollars") %>%
  filter(race =="All Races") %>%
  group_by(year) %>%
  mutate(pct = round(income_dollars/sum(income_dollars),2))
summary(im2$year)
```

```{r}
# avg household income
h1 = ggplot(im2, aes(year, income_dollars, fill=fct_rev(income_quintile), label=income_quintile)) + 
  geom_stream(type="ridge", bw=0.75) + 
  geom_stream_label(size = 3, type = "ridge", color="black", family="Courier Bold") +
  scale_fill_manual(values = c("#83c5be","#cbf3f0","#f4f3ee","#ffbf69","#ff9f1c")) + 
  theme_light() +
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        plot.title = element_text(face="bold",size=11),
        plot.subtitle= element_text(size=9.5),
        plot.caption= element_text(hjust=0.5, color="slategrey", size=8)
        ) + 
  scale_x_continuous(breaks=seq(1970,2015,5)) + 
  scale_y_continuous(labels=dollar_format()) +
  coord_cartesian(expand=FALSE) + 
  labs(x="",
       y="",
       title="Average household income, by quintile in 2019 dollars (1976 to 2019)",
       subtitle="",
       caption= "Data from Urban Institute and US Census")
```


```{r}
# quintile share
h2 = ggplot(im2, aes(year, pct, fill=fct_rev(income_quintile), label=income_quintile)) + 
  geom_stream(type="proportional", bw=0.5) + 
  geom_stream_label(size = 3, type = "proportional", color="black", family="Courier Bold") +
  scale_fill_manual(values = c("#83c5be","#cbf3f0","#f4f3ee","#ffbf69","#ff9f1c")) + 
  theme_light() +
  theme(legend.position="none",
        panel.grid.minor = element_blank(),
        plot.title = element_text(face="bold",size=11),
        plot.subtitle= element_text(size=9.5),
        plot.caption= element_text(hjust=0.5, color="slategrey", size=8)
        ) +  
  scale_y_continuous(labels=percent_format(accuracy=1), breaks=seq(0,1,0.2)) + 
  scale_x_continuous(breaks=seq(1970,2015,5)) + 
  coord_cartesian(expand=FALSE) + 
  labs(x="",
       y="",
       title="Share of aggregate household income, by quintile (1976 to 2019)",
       subtitle="",
       caption= "Data from Urban Institute and US Census")
```

```{r, fig.height=2.7, fig.width=6}
# combine both plots
grid.arrange(h1, h2, ncol=2)
```

### P8: retirement (average by racial group)

```{r, warning=FALSE, message=FALSE}
retirement2 = retirement %>% 
  mutate(ret= round(retirement/1000)) %>%
  filter(year>2000)
retirement2$year=as.factor(retirement2$year)

newggslopegraph(dataframe = retirement2,
                Times = year,
                Measurement = ret,
                Grouping = race,
                LineThickness = 1,
                LineColor = c("White"="#fb8500","Hispanic"="#006d77", "Black"="#9e2a2b"),
                Title = "Average family liquid retirement savings by racial group",
                SubTitle = "Retirement dollars (in thousands) from 2001 to 2016, normalized to 2016 dollar ",
                Caption = "Data from Urban Institute and US Census",
                YTextSize = 3.4,
                DataTextSize = 3,
                DataLabelFillColor = "white",
                DataLabelPadding=0.05,
                CaptionJustify = "center"
                )
```


### P9: income mean
* reference: [TidyTuesday's cleaning script](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-02-09/readme.md)

```{r}
summary(income_mean$year)
```


```{r}
im = income_mean %>%
  filter(income_quintile != "Top 5%") %>%
  mutate(income_quintile = factor(income_quintile, levels = c("Lowest", "Second", "Middle", "Fourth", "Highest"))) %>%
  filter(dollar_type == "2019 Dollars") %>%
  filter(race %in% c("White Alone", "Black Alone", "Hispanic")) %>%
  ggplot(aes(x = year, y = income_dollars, color = race, fill = race, group = race)) +
  geom_line() +
  facet_wrap(~income_quintile, scales = "free_y") + 
  scale_color_manual(values=c("#9e2a2b","#006d77","#fb8500")) + 
  theme_light() +
  theme(strip.text = element_text(face="bold"),
        strip.background= element_rect(fill="#1d3557")) + 
  labs(color="Racial Group",
       title="Income Quintile",
       subtitle="Mean income received by each fifth of each racial group from 1967 to 2019",
       x="Income (in dollars)",
       y="Year", caption= "Data from Urban Institute and US Census")

reposition_legend(im, 'center', panel='panel-3-2')
```

### P10: income distribution
* reference: [TidyTuesday's cleaning script](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-02-09/readme.md)

```{r}
income_levels <- c(
  "Under $15,000",
  "$15,000 to $24,999",
  "$25,000 to $34,999",
  "$35,000 to $49,999",
  "$50,000 to $74,999",
  "$75,000 to $99,999",
  "$100,000 to $149,999",
  "$150,000 to $199,999",
  "$200,000 and over"
)


income_distribution %>%
  filter(race %in% c("Black Alone", "White Alone", "Hispanic (Any Race)")) %>%
  filter(year >= 1980) %>%
  mutate(income_bracket = factor(income_bracket, levels = income_levels)) %>%
  ggplot(aes(x = year, y = income_distribution, color = race, fill = race)) +
  geom_col(position = "fill") +
  theme_light() + 
  facet_wrap(~income_bracket) +
  scale_fill_manual(values=c("#ffc145","#c0c0c0","#3a6ea5")) + 
  scale_color_manual(values=c("#ffc145","#c0c0c0","#3a6ea5")) +
  coord_cartesian(expand=FALSE) + 
  theme(strip.text = element_text(face="bold", color="#273e47"),
        strip.background= element_rect(fill="white"),
        plot.caption = element_text(size=8, face="italic",hjust=0),
        plot.caption.position = "plot") + 
  labs(fill="Racial group", 
       color="Racial group",
       y="Income Distribution",
       x="Year",
       caption= "Data from Urban Institute and US Census")
```


### P11: ratios
#### Student debt to family income
* reference: [David Zoller -
Change in student loan debt to householdincome by race ](https://twitter.com/winterstat/status/1359289102739251203/photo/1) 

```{r}
summary(student_debt$year)
summary(income_distribution$year)
```

```{r}
# select data 
inc = income_distribution %>%
  filter(race %in% c("Black Alone", "White Alone", "Hispanic (Any Race)")) %>% 
  mutate(race = recode(race, "White Alone"="White", "Black Alone"="Black","Hispanic (Any Race)"="Hispanic")) %>% 
  filter(year>=1989) %>%
  filter(year<=2016) %>%
  select(year, race, income_mean) %>%
  distinct(year, race, income_mean, .keep_all=TRUE) #drop dups

stu = student_debt %>% select(-loan_debt_pct)

# merge 
joined = inner_join(stu, inc, by=c("year","race")) %>% 
  mutate(ratio = round(loan_debt/income_mean,3)) #get ratio
head(joined)
```

```{r}
left_plot = joined %>% 
  ggplot(aes(y=year, x=ratio)) + 
  geom_line(aes(group=year), color="#1d3557",alpha=0.8) +
  geom_point(aes(color=race), size=2.7, alpha=0.8) + 
  scale_x_continuous(limits=c(0,0.253),labels=percent_format(accuracy=1)) + 
  scale_y_continuous(limits=c(1988, 2019),breaks=seq(1989, 2016,3)) + 
  theme_minimal() + 
  theme(legend.position="none",
        panel.grid.minor=element_blank(),
        panel.grid.major.y=element_blank(),
        plot.title=element_text(face="bold",size=14), 
        plot.title.position = "plot",
        plot.subtitle=element_text(size=9.5, color="#1d3557"),
        plot.caption =element_text(hjust=0.5, color="#333533",size=8)) + 
  scale_color_manual(values=c("#9e2a2b","#006d77","#fb8500")) +
  coord_cartesian(expand=FALSE) +
  labs(x="",
       y="",
       color="",
       caption="Data from Urban Institute and US Census",
       title="Student loan debt to household income, by race",
       subtitle="Average student loan debt as a percentage of average yearly income")  +
  geom_text(data=joined,aes(y=2017.5, x=0.1, label="Hispanic"), color="#006d77", size=3) +
  geom_text(data=joined,aes(y=2017.5, x=0.12, label="White"), color="#fb8500", size=3) + 
  geom_text(data=joined,aes(y=2017.5, x=0.23, label="Black"), color="#9e2a2b", size=3) 
  
```


#### Liquid retirement to family wealth
* reference: [Sonia Winter - Family retirement savings by race in the USA](https://twitter.com/winterstat/status/1359289102739251203/photo/1) 

```{r}
# join retirement and race_wealth
wr = race_wealth %>%
  filter(type=="Average" & race !="Non-White") %>%
  select(-type) %>%
  full_join(retirement, by=c("year","race")) %>%
  mutate(pct = retirement/wealth_family) # ratio variable 

# liquid retirement as a pct of overall wealth
right_plot = wr %>% filter(year >= 1989) %>%
  ggplot(aes(y = year, x = pct)) +
  geom_line(aes(group = year), color="#1d3557",alpha=0.8) +
  geom_point(aes(color = race), size=2.7, alpha=0.8, show.legend = FALSE) +
  scale_x_continuous(limits=c(0,0.253),labels=percent_format(accuracy=1)) + 
  scale_y_continuous(limits=c(1988, 2019),breaks=seq(1989, 2016,3)) +
  scale_color_manual(values=c("#9e2a2b","#006d77","#fb8500")) +
  labs(title= "Family liquid retirement to family wealth, by race",
       subtitle = "Average liquid retirement as a percentage of average family wealth, normalized to 2016 dollars", 
       caption="Data from Urban Institute",
       x = "", y = "") +
  theme_minimal() + 
  theme(legend.position="none",
        panel.grid.minor=element_blank(),
        panel.grid.major.y=element_blank(),
        plot.title=element_text(face="bold",size=14), 
        plot.title.position = "plot",
        plot.subtitle=element_text(size=9.5, color="#1d3557"),
        plot.caption =element_text(hjust=0.5, color="#333533",size=8)) + 
  geom_text(data=joined,aes(y=2017.5, x=0.14, label="Hispanic"), color="#006d77", size=3) +
  geom_text(data=joined,aes(y=2017.5, x=0.165, label="White"), color="#fb8500", size=3) + 
  geom_text(data=joined,aes(y=2017.5, x=0.185, label="Black"), color="#9e2a2b", size=3) + 
  coord_cartesian(expand=FALSE) 
```

```{r, fig.height=2.7, fig.width=6}
# combine both plots
grid.arrange(left_plot, right_plot, ncol=2)
```




