---
title: "Tidy Tuesday Week 29/2021"
output: html_notebook
---

[TidyTuesday](https://github.com/rfordatascience/tidytuesday) week 29 [Scooby Doo Episodes](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-07-13/readme.md), data from [Kaggle](https://www.kaggle.com/williamschooleman/scoobydoo-complete) and aggregated by  [plummye](https://www.kaggle.com/williamschooleman).


```{r}
# load libaries
library(tidyverse)
library(scales)
library(lubridate)
library(skimr)
library(tidytext)
library(ggtext)
library(ggpubr)
library(colorspace)
library(ggsci)
library(viridis)
library(gggrid) 
library(packcircles)
library(ggmosaic)
library(ggbump)
library(ggsankey)
```

```{r, message=F}
# import data
scoobydoo <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-13/scoobydoo.csv')

# import data (change NULL to NA)
scoobydoo1 <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-13/scoobydoo.csv', na="NULL")
```

```{r}
# summary

# numeric variables
scoobydoo1 %>% 
  select_if(is.numeric) %>%
  skim()

# logical variables
scoobydoo1 %>% 
  select_if(is.logical) %>%
  skim()

# character variables
scoobydoo1 %>% 
  select_if(is.character) %>%
  mutate_if(is.character,as.factor) %>%
  skim()
```


#### Main character action count in TV series
* shared on [Twitter](https://twitter.com/leeolney3/status/1414843982874648577)

```{r}
# tv series: action count by character
table1 = scoobydoo %>% 
  filter(format=="TV Series") %>%
  select(index,caught_fred:snack_scooby) %>%
  pivot_longer(!index) %>%
  extract(name, c("type", "char"), "(.*)_(.*)") %>%
  filter(value=="TRUE") %>%
  group_by(type, char) %>% 
  tally() 

p1 = table1 %>% 
  mutate(char=str_to_title(char)) %>%
  ggplot(aes(x=type, y=n, fill=type)) + 
  geom_bar(stat="identity",width=0.9, alpha=0.95) + 
  facet_wrap(~char, ncol=5) + 
  scale_fill_manual(values=c("#76a2ca","#cd7e05","#b2bb1b","#7c68ae")) +
  scale_x_discrete(expand=c(.3,.3)) +
  theme_minimal() + 
  theme(panel.grid.major.x=element_blank(),
        panel.grid.minor.y=element_blank(),
        plot.subtitle=element_markdown(size=10, face="bold", hjust=0.5),
        axis.text.x=element_blank(),
        strip.text=element_text(face="bold", size=8),
        legend.position="none",
        plot.title.position = "plot",
        plot.title=element_text(hjust=0.5, face="bold",size=12),
        axis.title=element_blank(),
        #plot.background = element_rect(fill="#f5f7f5", color=NA),
        plot.margin=unit(c(1,1,0.5,1),"cm"),
        axis.text.y=element_text(size=7),
        panel.grid=element_line(color="#bccbc0", linetype = 3)) + 
  labs(subtitle= "<span style = 'color:#76a2ca'><b>Captured</b></span>, 
       <span style = 'color:#cd7e05'><b>Caught</b></span>,
       <span style = 'color:#b2bb1b'><b>Snack eaten</b></span>, and
       <span style = 'color:#7c68ae'><b>Unmask</b></span><br>",
       title="Scooby Doo TV Series")
```

```{r}
# tv series: captured vs caught
d1 = scoobydoo %>% 
  filter(format=="TV Series") %>%
  select(index,caught_fred:captured_scooby) %>%
  pivot_longer(!index) %>%
  extract(name, c("type", "char"), "(.*)_(.*)") %>%
  filter(value=="TRUE") %>%
  group_by(type, char) %>% 
  tally() %>%
  ungroup() 

d2 = d1 %>% 
  group_by(char) %>%
  mutate(diff= n-lag(n),
         diff2=ifelse(is.na(diff),mean(diff,na.rm=T),diff),
         cat = ifelse(diff2>0,"Caught > Captured","Captured > Caught")) %>%
  arrange(char, type) %>%
  select(-diff) %>%
  ungroup() %>%
  mutate(char=str_to_title(char),
         type=str_to_title(type))

d3 = d2 %>% group_by(char) %>% mutate(mea= median(n)) %>% filter(type=="Caught")

p2 = d2 %>%
  ggplot(aes(y=fct_rev(char), x=n)) + 
  geom_point(aes(shape=type),size=3) + 
  geom_line(aes(group=char, color=cat)) + 
  scale_shape_manual(values=c(1,19)) + 
  scale_color_manual(values =c("#76a2ca","#cd7e05")) + 
  scale_x_continuous(limits=c(0,120), breaks=seq(30,120,20)) +
  geom_text(data = d3, 
            aes(x=mea, y= char, label=abs(diff2)),size=3, color="black", vjust=-0.9) + 
  geom_text(data =d3, aes(x=0, y=char, label=char, color=cat), size=4, hjust=0, fontface="bold",
            show.legend = F) + 
  theme(legend.position = "top",
        axis.text.y=element_blank(),
        axis.title=element_blank(),
        plot.margin=unit(c(1,1,0.5,1),"cm"),
        plot.title.position = "plot",
        plot.title=element_text(face="bold", hjust=0.5,size=12),
        panel.grid.minor=element_blank()) + 
  labs(x="Count", y="Character", shape="", color="",
       title="Scooby Doo TV Series: Captured vs. Caught",
       caption="\nTidy Tuesday Week 29 | Data from Kaggle")  + 
  guides(shape = guide_legend(order = 1),color = guide_legend(order = 2, override.aes = list(size = 2)))
```

```{r,fig.width=4, fig.height=3.5}
# combine p1 and p2
ggarrange(p1,p2,nrow=2)
```

ALT text: Bar plot and dot plot showing the action count of main characters in Scooby Doo TV series. Bar plot shows count of captured, caught, snack eaten and unmask by main characters, where Daphnie has more snacks eaten than caught. Dot plot shows the count of captured vs. caught, where Fred and Scooby has more caught than captured while the others (Daphnie, Shaggy and Velma) has more captured than caught. 



#### Proportion of real monsters by format

```{r}
# packing circles chart

data2a = scoobydoo %>% 
  group_by(format,monster_real) %>% tally() %>% 
  mutate(prop=n/sum(n),prop = round(prop*100)) %>% ungroup()

data2b = data2a %>% group_by(format) %>% count(wt=prop) %>% ungroup()

data2c = data2b %>% 
  pmap_df(
  .f = ~circleProgressiveLayout(rep(0.5, ..2))) %>% 
  mutate(format = rep(data2a$format, data2a$prop),
         monster_real = rep(data2a$monster_real, data2a$prop)
  )

# graphic params
my_gpar <- gpar(
  col = "black",
  fontsize = 7
)

# plot
p3 =plot2 = data2c%>% 
    ggplot(aes(x, y, fill = monster_real)) + 
    geom_point(size = 2.75,pch = 21, color="white") + 
  facet_wrap(~format) + 
  scale_fill_manual(values=c("#128a84","#D0D61B","#F7921E")) +
  scale_y_continuous(expand=c(.1,.1)) +
  scale_x_continuous(expand=c(.2,.2)) +
  theme_void() +
  theme(legend.position="top",
        plot.title=element_text(hjust=0.5, face="bold", size=12),
        plot.subtitle=element_text(hjust=0.5, size=8),
        strip.text.x=element_text(size=8,face="bold", margin=margin(b=5,t=5)),
        plot.caption=element_text(size=8),
        legend.text = element_text(size=8),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        legend.margin=margin(t = 0, unit='cm'),
        legend.box.margin=margin(t=-10, b=0)) +
  guides(fill=guide_legend(ncol=3,title.position = "top", 
                           title.hjust = .5, override.aes = list(size=3))) + 
  labs(subtitle="(Proportion of real monsters, where each dot represents 1%)", fill="",
       title="Real Monsters in Scooby Doo, by Format")
p3
```

#### Proportion of monster type 

```{r}
# proportion of monster type
scoobydoo1 %>% separate_rows(monster_type,sep=",", convert=T) %>%
  select(index, monster_type) %>%
  drop_na(monster_type) %>%
  mutate(monster_type=str_trim(monster_type),
         monster_type=recode(monster_type, 
                             Disguised = "Disguise",
                             Disguised = "Disugised",
                             Possessed="Possessed Object"
                             )) %>%
  #count(monster_type, sort=T) %>%
  filter(monster_type!="NULL") %>%
  mutate(mt = fct_lump(monster_type, 10)) %>%
  count(mt, sort=T) %>%
  mutate(perc= paste0(sprintf("%2.1f", n/sum(n)*100),"%"),
         mt = fct_rev(fct_inorder(mt)),
         mt = fct_relevel(mt, "Other",after=0),
         col=case_when(row_number()==1 ~ "#76a2ca", 
                         row_number()==2 ~ "#cd7e05",
                         row_number()==3 ~ "#b2bb1b",
                         row_number()==4 ~ "#7c68ae",
                         mt =="Other" ~"grey70",
         TRUE~"grey55")) %>%
  ggplot(aes(y=mt, x=n, fill=col)) + 
  geom_col(width=0.7, show.legend = F, alpha=0.95) +
  scale_fill_identity() + 
  scale_color_identity() + 
  geom_text(aes(label=perc), size=3, color="white", hjust=1.3,fontface="bold") + 
  geom_text(aes(x=-5, label=mt, color=col),size=3.5, hjust=1, fontface="bold") + 
  scale_x_continuous(expand=expansion(mult=c(.3,.1))) +
  theme_void() + 
  theme(plot.subtitle=element_text(hjust=0.06),
        plot.margin=unit(c(1,1,1,1),"cm")) +
  labs(subtitle="Proportion of monster types in Scooby Doo\n")
```



#### Proportion of monster gender by format

```{r}
# monster gender
scoobydoo1 %>% select(index, format, monster_gender) %>%
  drop_na(format, monster_gender) %>%
  separate_rows(monster_gender,sep=",", convert=T) %>%
  filter(monster_gender!="",monster_gender!="None") %>%
  group_by(format, monster_gender) %>% tally() %>%
  mutate(prop=round(n/sum(n),3)) %>%
  ggplot(aes(y=fct_rev(format), x=prop, fill=monster_gender)) + 
  geom_col(position=position_dodge2(width=0.2,preserve = "single"), width=0.5,alpha=0.8) + 
  scale_fill_manual(values=c("#ee9b00","#005f73")) +
  scale_x_continuous(labels=scales::percent_format(),position="top") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10)) + 
  theme_minimal() +
  theme(legend.position="top",
        panel.grid.minor=element_blank(),
        axis.title=element_blank(),
        legend.justification = "left",
        legend.margin = margin(l=15),
        plot.title.position = "plot",
        axis.text.x.top = element_text(margin = margin(b = -10, unit = "pt")),
        axis.text.y.left = element_text(margin = margin(r = -10, unit = "pt")),
        plot.margin=unit(c(1,1,1,1),"cm")) + 
  guides(fill=guide_legend(keyheight = 1, keywidth = 0.4, reverse=T)) + 
  labs(fill="",
       subtitle="Proportion of monster gender by format") 
```


```{r}
# percentage of female monsters by monster amount
gender_amount = scoobydoo1 %>% 
  drop_na(monster_gender,monster_amount) %>%
  #mutate(ma = fct_lump(factor(monster_amount),5, other_level = "5 and above")) %>%
  select(index, monster_amount, monster_gender) %>%
  separate_rows(monster_gender,sep=",", convert=T) %>%
  group_by(index,monster_amount) %>% 
  count(monster_gender) %>%
  ungroup() %>%
  filter(monster_gender!="",
         monster_gender!="None") %>%
  group_by(monster_amount, monster_gender) %>% tally(n) %>%
  mutate(prop=round(n/sum(n),3))
  
gender_amount %>% filter(monster_gender=="Female") %>% arrange(desc(prop))
```

#### Monster gender and amount

```{r}
# mosiac plot: monster gender and monster amount
x= scoobydoo1 %>% 
  drop_na(monster_gender,monster_amount) %>%
  #mutate(ma = fct_lump(factor(monster_amount),4, other_level = "5 and above")) %>%
  select(index, monster_amount, monster_gender) %>%
  separate_rows(monster_gender,sep=",", convert=T) %>%
  filter(monster_gender!="",
         monster_gender!="None")

ggplot(data = x) + geom_mosaic(aes(x=product(monster_gender,monster_amount), fill=monster_gender)) + 
  scale_fill_manual(values=c("#ee9b00","#005f73")) + 
  theme(panel.grid=element_line(size=.3),
        plot.margin=unit(c(1,2,1,2),"cm"),
        legend.position = "none",
        axis.title=element_text(size = 9),
        plot.title.position = "plot") + 
  labs(x="Monster amount", y="Monster gender",
       subtitle="Monster amount and gender")
```

#### Monster count, imdb score, format

```{r, warning=F}
# monster count, imdb score, format
scoobydoo %>% 
  mutate(imdb2 = parse_number(imdb)) %>%
  drop_na(imdb2) %>%
  filter(format!="Movie (Theatrical)") %>%
  ggplot(aes(x=monster_amount, y=imdb2, color=format)) + 
  geom_point(alpha=0.5) +
  scale_color_npg() + 
  theme_light(base_size=10) +
  theme(legend.position = "none",
        panel.grid.minor=element_blank(),
        panel.grid.major=element_line(size=.3),
        strip.text=element_text(size = 9),
        strip.background = element_rect(fill="slategrey")) + 
  facet_wrap(~format) + 
  labs(x="Monster count", y="IMDB rating",
       subtitle = "Monster count and IMDB rating, by format")
```

#### TV series: monster_amount and imdb score

```{r}
# TV series: monster_amount and imdb score 
scoobydoo1 %>% drop_na(imdb,monster_amount) %>%
  mutate(ma = fct_lump(factor(monster_amount),5, other_level = "5 and above")) %>%
  filter(format=="TV Series") %>%
  ggplot(aes(x=ma, y=imdb,color=ma)) + 
  geom_half_boxplot(outlier.size=-1) + 
  geom_half_point(alpha=0.7) + 
  theme(legend.position = "none",
        plot.margin=unit(c(1,2,1,1),"cm"),
        panel.grid.minor=element_blank(),
        axis.text=element_text(size=9),
        axis.title.x=element_text(margin=margin(t=10),size=9),
        axis.title.y = element_text(margin=margin(r=10),size=9),
        plot.title.position = "plot") + 
  scale_color_npg() + 
  labs(x= "Monster amount", y="IMDB score",
       subtitle="IMDB score by monster amount\n")
```

#### Monster amount and network

```{r, warning=F, message=F}
# monster amount ny network
scoobydoo %>% 
  mutate(network2 = fct_lump(network,7)) %>%
  ggplot(aes(y=fct_rev(network2), x=monster_amount, color=network2, fill=network2)) + 
  geom_density_ridges(
    jittered_points = TRUE,
    position = position_points_jitter(width = 0.05, height = 0),
    point_shape = '|', point_size = 3, point_alpha = 0.7, alpha = 0.3,
  ) +
  scale_fill_d3() +
  scale_color_d3() +
  scale_y_discrete(expand=expansion(mult=c(0.1,0.25))) +
  theme(panel.grid.minor=element_blank(),
        legend.position="none",
        plot.title.position = "plot",
        axis.title=element_text(size=9)) + 
  labs(y="Network", x="Monster amount", subtitle="Monster amount by network")
```

### Date aired and snacks rank
* reference: https://twitter.com/isitmanu/status/1415058405811646465/photo/1

```{r}
# most unmask by
as_decade <- function(year) {
  return(round(year / 10) * 10)
}

`%not_in%` <- Negate(`%in%`)

unmask_rank <-
  scoobydoo1 %>%
  select(starts_with('unmask'), date_aired) %>%
  mutate(across(where(is.character),
                as.logical)) %>%
  mutate(year = year(date_aired),
         decade = as_decade(year)) %>%
  pivot_longer(where(is.logical),
               names_to = 'unmask_by',
               names_prefix = 'unmask_') %>%
  mutate(unmask_by = str_to_title(unmask_by)) %>%
  filter(unmask_by %not_in% c('Other', 'Not')) %>%
  count(unmask_by, decade, wt = value) %>%
  arrange(desc(n)) %>%
  group_by(decade) %>%
  mutate(rank = rank(n, ties.method = 'first')) %>%
  ungroup()

unmask_rank %>%
  ggplot(aes(x = decade, y = rank, color = unmask_by)) +
  geom_bump(smooth = 8) +
  geom_point() +
  geom_text(data =  unmask_rank %>% filter(decade == min(decade)),
            aes(x = decade - 1, label = unmask_by), size = 3.5, hjust = 1, fontface = 'bold')  +
  geom_text(data =  unmask_rank %>% filter(decade == max(decade)),
            aes(x = decade + 1, label = unmask_by), size = 3.5, hjust = 0,fontface = 'bold') +
  geom_bump(size = 1.5, smooth = 8) +
  geom_point(size = 3) +
  #scale_y_reverse(breaks = seq(1, 5, 1), expand=c(0.05,0.05)) +
  #scale_y_discrete(breaks = seq(1, 5, 1), expand=c(0.05,0.05)) +
  scale_x_continuous(limits = c(1965, 2025),breaks = seq(1970, 2020, 10), expand=c(0.05,0.05)) + 
  scale_color_manual(values=c("#966a00","#79af30","#622486","#F7801E","#76a2ca")) +
  scale_shape_manual(values=c(21:25)) +
  theme_light(base_size = 10) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        panel.border=element_blank(),
        axis.title=element_blank(),
        plot.title.position="plot",
        plot.margin=unit(c(1.5,2,1.5,2),"cm")
        ) + 
  labs(subtitle="Characters with most unmask by decade\n")
```

### Date aired, cumulative snacks, cumulative caught
* reference: https://twitter.com/etmckinley/status/1415015967957176329/photo/1

```{r,fig.width=4, fig.height=2}
# date aired
theme_set(theme_minimal(base_size = 10)) 
theme_update(legend.position = "none", 
        axis.title = element_text(size=9),
        plot.margin=unit(c(0.5,0.5,0.5,0.5),"cm"),
        plot.title.position = "plot",
        panel.grid.minor=element_blank())

# cumulative snack (date_aired)
snack_tab = scoobydoo %>%
  select(index,
         series_name,
         season,
         title,
         date_aired,
         starts_with("snack")) %>%
  pivot_longer(starts_with("snack")) %>%
  group_by(name) %>%
  arrange(date_aired) %>%
  mutate(cumulative = ifelse(value == TRUE, 1, 0),
         cumulative = cumsum(cumulative),
         name = str_to_title(gsub("snack_", "", name)),
         name=case_when(name == "Daphnie"~"Daphne",
                        TRUE ~ name))

plot1 = snack_tab %>% 
  arrange(date_aired) %>% 
  ggplot(aes(x=date_aired, y=cumulative))+
  geom_path(aes(color=name), size=1)+
  geom_text(data=. %>% group_by(name) %>% arrange(date_aired) %>% slice(n()) ,
                  aes(color=name, label=name), hjust=0, nudge_x = 100, size=3, fontface="bold")+
  scale_x_date(limits=c(min(unmask$date_aired), max(unmask$date_aired)+3000)) + 
  scale_color_manual(values=c("#966a00","#79af30","#622486","#F7801E","#76a2ca")) +
  labs(x="Date aired", y="Cumulative", subtitle="Snack eaten by\n")


# cumulative snack (date_aired)
caught_tab = scoobydoo %>%
  select(index,
         series_name,
         season,
         title,
         date_aired,
         starts_with("caught")) %>%
  select(-caught_not,-caught_other) %>%
  pivot_longer(starts_with("caught")) %>%
  group_by(name) %>%
  arrange(date_aired) %>%
  mutate(cumulative = ifelse(value == TRUE, 1, 0),
         cumulative = cumsum(cumulative),
         name = str_to_title(gsub("caught_", "", name)),
         name=case_when(name == "Daphnie"~"Daphne",
                        TRUE ~ name))

plot2 = caught_tab %>%
  arrange(date_aired) %>% 
  ggplot(aes(x=date_aired, y=cumulative))+
  geom_path(aes(color=name), size=1)+
  geom_text(data=. %>% group_by(name) %>% arrange(date_aired) %>% slice(n()) ,
                  aes(color=name, label=name), hjust=0, nudge_x = 100, size=3, fontface="bold")+
  scale_x_date(limits=c(min(unmask$date_aired), max(unmask$date_aired)+3000)) + 
  scale_color_manual(values=c("#966a00","#79af30","#622486","#F7801E","#76a2ca")) +
  labs(x="Date aired", y="Cumulative",subtitle="Caught by\n")

ggarrange(plot1,plot2, ncol=2)
```


#### Sankey chart
* reference: https://twitter.com/allisonkoh_/status/1414925075439099909/photo/1

```{r}
stab <- scoobydoo %>% 
  filter(format != c("Crossover", "Movie", "Movie (Theatrical)")) %>% 
  select(
    c("index"), 
    starts_with(c("caught", "captured", "unmask","snack")), 
    -ends_with(c("other", "not"))) %>% 
  pivot_longer(
    cols = starts_with(c("caught", "captured", "unmask","snack")),
    names_to = c("action", "character"), 
    names_sep = "_",
    values_to = "value") %>% 
  mutate(character=case_when(character == "daphnie"~"daphne",
                        TRUE ~ character)) %>%
  mutate(
    value = as.integer(as.logical(value)), 
    action = recode(
      action, 
      unmask = "unmasked\nmonster",
      caught = "caught\nmonster",
      captured = "was captured",
      snack = "snack"),
    character = unlist(TC(character))) %>% 
  filter(value != 0) 

stab_long<- stab  %>% 
  make_long(character, action) 
```


```{r}
stab_long %>% 
  ggplot(aes(x = x, next_x = next_x, node = node, next_node = next_node, fill = factor(node), label = node)) + 
  geom_sankey(flow.alpha = .6) +
  geom_sankey_label(size = 3, color = "black", fill = "white") + 
  scale_fill_manual(values=c("grey20","#ea7317","#119da4","#ffc857","#4b3f72",
                             "grey40","grey60","#90be6d","grey")) + 
  theme_void() + 
  theme(legend.position = "none")
```


```{r}
# amount
amount = scoobydoo %>% 
  filter(format=="TV Series") %>%
  select(date_aired, suspects_amount, culprit_amount, monster_amount) %>%
  mutate(Suspect = cumsum(suspects_amount),
         Culprit = cumsum(culprit_amount),
         Monster = cumsum(monster_amount)) %>%
  select(-suspects_amount,-culprit_amount, -monster_amount) %>%
  pivot_longer(!date_aired) 

amount %>%
  ggplot(aes(x=date_aired, y=value)) + 
  geom_path(aes(color=name), size=1) + 
  geom_text(data=. %>% group_by(name) %>% arrange(date_aired) %>% slice(n()) ,
                  aes(color=name, label=name), hjust=0, nudge_x = 100, size=3.5, fontface="bold")+
  scale_color_manual(values=c("#b2bb1b","#cd7e05","#76a2ca")) + 
  scale_x_date(limits=c(min(unmask$date_aired), max(unmask$date_aired)+3000)) + 
  theme(plot.margin=unit(c(1,2,1,2),"cm")) +
  labs(x="Date aired", y="Cumulative amount", subtitle="Scooby Doo TV Series (1969 to 2021)\nCumulative suspect, monster and culprit amount\n")
```

#### Phrase, monster subtype, motive
```{r}
# phrase count
scoobydoo1 %>% 
  select(index,jeepers:rooby_rooby_roo) %>%
  pivot_longer(!index) %>%
  filter(value!=0) %>%
  mutate(phrase= str_replace_all(name, "_"," "),
         phrase = str_to_title(phrase)) %>%
  group_by(phrase) %>% tally(value, sort=T) 

# monster subtype count
scoobydoo1 %>% 
  separate_rows(monster_subtype,sep=",", convert=T) %>%
  drop_na(monster_subtype) %>%
  count(monster_subtype, sort=T)

# motive count
scoobydoo1 %>% 
  count(motive, sort=T)
```