---
title: "Makeup Shades"
date: "2021-03-30"
output: html_notebook
---

#### Data visualization exercise

This notebook uses [TidyTuesday](https://github.com/rfordatascience/tidytuesday/) week 14 [Makeup Shades](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-30/readme.md), data from [The Pudding data](https://github.com/the-pudding/data/tree/master/foundation-names). And the corresponding article can be found at 
[The Pudding](https://pudding.cool/2021/03/foundation-names/). 



```{r, warning=FALSE, message=FALSE}
# load libraries
library(tidyverse)
library(scales)
library(ggtext)
library(patchwork)

# set theme
theme_set(theme_minimal(10))
```

```{r}
# import data
sephora <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/sephora.csv')
ulta <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/ulta.csv')
allCategories <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allCategories.csv')
allShades <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allShades.csv')
allNumbers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-30/allNumbers.csv')
```

```{r}
head(allShades)
```

#### Boxplot: Foundation lightness of brands with more than 5 foundation products
* shared on [Twitter](https://twitter.com/leeolney3/status/1376644323991818247/photo/1)
* alt text: Boxplot showing the makeup foundation shades of fifteen brands that have more than 5 foundation products, where the brand Too Faced has the darkest foundation shade.

```{r}
allShades %>% group_by(brand) %>% summarise(product_n=n_distinct(product)) %>% arrange(desc(product_n))
```

```{r}
# brands with >5 products
brand_prod = allShades %>% group_by(brand) %>% summarise(product_n=n_distinct(product)) %>% arrange(desc(product_n)) %>% filter(product_n>5)
brand_names = brand_prod$brand 
brand_names
```


```{r, fig.height=3.2, fig.width=4}
allShades %>% 
  filter(brand %in% brand_names) %>%
  ggplot(aes(y=fct_reorder(brand,lightness,.fun = median, .desc=TRUE), x=lightness)) + 
  geom_boxplot() + 
  theme_minimal(base_size = 10) + 
  theme(panel.grid.minor.x=element_blank(),
        plot.title.position = "plot",
        plot.caption.position = "plot",
        plot.title=element_text(face="bold",size=15),
        axis.title=element_text(face="bold"),
        plot.caption=element_text(hjust=0),
        plot.margin=margin(1,1,1,1,"cm")) + 
  labs(x="Lightness", y="Brand",
       caption="\nTidy Tuesday Week 14 | Data from The Pudding",
       title="Foundation lightness of brands with more than 5 foundation products",
       subtitle = "Lightness represented as a decimal from 0 to 1, where 0 is pure black and 1 is pure white\n") 
```
```{r, fig.height=3.2, fig.width=4, warning=FALSE, message=FALSE}
# dot plot
allShades %>% 
  filter(brand %in% brand_names) %>%
  ggplot(aes(y=lightness, x=fct_reorder(brand,lightness,.fun = median, .desc=TRUE))) + 
  geom_dotplot(fill="slategrey",
               binaxis = "y",
               binwidth = 0.005,
               stackdir="center", show.legend = F, size=1, color=NA, alpha=0.4) +
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                geom = "pointrange", width = 0.5, size=0.2, alpha=0.9, color="#f77f00") +
  coord_flip() + 
  theme_minimal(base_size = 10) + 
  theme(
    #plot.margin=margin(1,1,1,1,"cm"),
    axis.title=element_text(face="bold"),
    plot.title.position = "plot",
    plot.subtitle=element_markdown()) +
  labs(y="",x="", title="Foundation lightness by brand",
       subtitle = "Foundation lightness by brand, from lowest to highest <span style = 'color:#f77f00'><b>Median</b></span> lightness\n")

```


```{r}
# products
# lightToDark: Whether this product line organizes their colors from light to dark (Note: a value of NA indicates that a product uses a number-based naming system, but not a sequential numbering system)

# number of distinct brands
n_distinct(allNumbers$brand)
# lightToDark table
allNumbers %>% group_by(lightToDark) %>% tally() %>% mutate("%"=round(n/sum(n)*100,2)) #table
```
```{r, warning=FALSE, messsage=FALSE}
# histogram: lightness
hist = allCategories %>%
  ggplot(aes(lightness)) +
  geom_histogram(alpha = 0.7, bins = 100, fill="slategrey") +
  xlim(0, 1) + 
  labs(title="Histogram: lightness") + 
  theme(plot.title.position = "plot")

# density plot: lightness and lightToDark
dens = allNumbers %>%
  ggplot(aes(lightness, color=lightToDark)) +
  geom_density(alpha = 0.6, fill=NA) +
  xlim(0, 1) + 
  scale_color_manual(values=c("#577590","#90be6d"), na.value="#f94144") + 
  labs(title="Density plot: lightness and lightToDark") + 
  theme(plot.title.position = "plot")

# plot
(hist/dens)
```

```{r}
# brands that uses sequential numbering system (SNS)
seq = allNumbers %>% 
  filter(!is.na(lightToDark)) %>% 
  group_by(brand,product,lightToDark) %>%
  tally() %>%
  ungroup() %>%
  group_by(brand,lightToDark) %>% 
  summarise(product_n=n_distinct(product))

# brands that uses SNS that have both classes of lightToDark
seq %>% group_by(brand) %>% summarise(class_n=n_distinct(lightToDark)) %>% filter(class_n==2) %>% count() %>% unlist() #no brands uses both classes

# brand and lightToDark
seq %>% group_by(lightToDark) %>% tally()
```

* 61 out of 64 makeup brands in the dataset uses sequential numbering system, of which 57 organizes their colors from light to dark




#### Average product shades by brand
```{r}
# number of shades by product
shades = allShades %>% 
  group_by(brand, product) %>% 
  summarise(shades_n = n_distinct(hex))

summary(shades$shades_n)

# products with more than 50 shades
shades %>% filter(shades_n>=50) %>% arrange(desc(shades_n))
```


```{r}
# brands, product count and average number of shades
shades2 = shades %>% 
  group_by(brand) %>%
  summarise(product_count=n_distinct(product),
            avg_shades = (mean(shades_n))) %>%
  arrange(desc(avg_shades))
shades2
```

```{r}
hist(shades2$product_count)
```


```{r}
shades2 %>% mutate(cat = ifelse(product_count==1,"1 product",">1 product")) %>% 
  group_by(cat) %>%
  summarise(avg1 = mean(avg_shades))
```

```{r, warning=FALSE, message=FALSE}
# Brands with 1 product versus Brands with >1 product
shades2 %>% mutate(cat = ifelse(product_count==1,"1 product",">1 product")) %>%
  ggplot(aes(x=avg_shades, y= cat)) + 
  geom_bar(stat="summary",fun.x="mean", width=0.5, alpha=0.5, fill="#b8b8d1") +
  geom_boxplot(fill=NA, width=0.2, outlier.size=-1, color="#f28f3b") +
  geom_jitter(height=0.2, size=1.5, alpha=0.8, color="#255f85") + 
  theme_minimal(base_size = 10) + 
  theme(panel.grid.minor.x=element_blank(),
        panel.grid.major.y=element_blank(),
        plot.margin=margin(1,1,1,1,"cm"),
        plot.title.position = "plot") + 
  labs(x="Average product shades (brand)",
       y="",
       title="Average product shades",
       subtitle="Brands with 1 product vs. Brands with >1 product")
```

#### 10 brands with the most shades

```{r}
# barplot of 10 brands with the most shades, with their palettes
# code reference: [Richard Vogg](https://t.co/y1IBB3fjMJ?amp=1)

brands_n = allCategories %>% 
  group_by(brand) %>%
  tally(sort=T) %>%
  slice(1:10) %>%
  .$brand # get brand names

brands_n

# sort 
sorted <- allCategories %>%
  group_by(brand) %>%
  arrange(brand,lightness) %>%
  mutate(rank=rank(lightness,ties.method = "first"))
```


```{r}
# function to create plot
get_brand_colors <- function(brand_name) {
  sorted %>%
    filter(brand==brand_name)
}

plot_brand_colors <- function(brand_data) {
  title <- brand_data[[1,1]]
  
  ggplot(brand_data,aes(x=rank,y=brand,fill=hex)) + geom_tile() +
    scale_fill_manual(values=brand_data$hex) +
    scale_x_continuous(limits=c(0,370)) +
    theme_minimal(base_size=10)+
    theme(legend.position = "none",
          axis.title = element_blank(),
          axis.text.x = element_blank(),
          axis.ticks = element_blank(),
          plot.subtitle = element_text(),
          panel.grid=element_blank()) + 
    coord_cartesian(expand=FALSE)
}

plots <- lapply(brands_n,function(x) {
  x %>% 
     get_brand_colors() %>% 
     plot_brand_colors()
  })
```


```{r}
# plot
(plots[[1]]) / (plots[[2]]) / (plots[[3]]) / (plots[[4]]) / (plots[[5]]) / (plots[[6]]) / (plots[[7]]) / (plots[[8]]) / (plots[[9]]) / (plots[[10]]) + 
  plot_annotation(title = "10 brands with the most foundation shades")
```



```{r}
# get 10 most frequent name (programmatically extracted word-based name of this particular shade) 
names1 = 
  allShades %>% 
  mutate(name=tolower(name)) %>%
  count(name) %>%
  drop_na() %>%
  arrange(desc(n)) %>%
  mutate(rank=rank(desc(n), ties.method = "first")) #code from [Cal Webb](https://t.co/KrE6NtQPaw?amp=1)

# join rank to data 
names2 = left_join(allShades %>% mutate(name = tolower(name)), names1, by = c("name")) %>% drop_na()

# shade colors
shade_col <- names2 %>% 
  filter(rank<=10) %>% pull(hex, hex) #pull hex code from [Jamie Avendano](https://t.co/hxmx55bxWC?amp=1)

# plot
names2 %>% filter(rank<=10) %>%
  ggplot(aes(y=fct_reorder(name,lightness,.fun = median, .desc=TRUE), x= lightness)) + 
  geom_boxplot(outlier.shape=NA, color="slategrey",width=0.7) + 
  geom_jitter(aes(color=hex),alpha=0.6, show.legend = F, height=0.2, size=1.5) + 
  scale_color_manual(values=shade_col) + 
  theme(plot.margin=margin(1,1,1,1,"cm"),
        plot.title.position = "plot"
        #plot.title=element_text(face="bold"),
        #axis.title=element_text(face="bold")
        ) + 
  labs(x="Name", y= "Lightness",
       title="Lightness of 10 Most Frequent Word-based Shade Name",
       subtitle="Where Lightness 0 is pure black and 1 is pure white\n")

```

