---
title: "Tidy Tuesday Week 26/2021"
output: html_notebook
---

[TidyTuesday](https://github.com/rfordatascience/tidytuesday) week 26 [Public Park Access](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-06-22/readme.md), data from [The Trust for Public Land](https://www.tpl.org/parks-and-an-equitable-recovery-parkscore-report).

```{r}
# Load libraries
library(tidyverse)
library(ggtext)
library(ggbump)
library(fmsb) 
library(ggsankey)
library(colorspace)
```

```{r}
# Import data
parks <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-22/parks.csv')
glimpse(parks)
```

```{r}
# Summary
summary(parks$year)
summary(parks$rank)
n_distinct(parks$city)
```


#### Normalised values (points)
* shared on [Twitter](https://twitter.com/leeolney3/status/1407205328249311235/photo/1)

```{r, warning=F}
# points
parks %>% select(year,
                 med_park_size_points,
                 park_pct_city_points,
                 pct_near_park_points,
                 basketball_points,
                 dogpark_points,
                 playground_points,
                 rec_sr_points,
                 restroom_points,
                 splashground_points,
                 amenities_points) %>%
  rename(`recreation_&_senior_center` = rec_sr_points,
         `median_park_size` = med_park_size_points,
         `parkland_of_city_area`=park_pct_city_points,
         `within_10_min_walk_to_park`=pct_near_park_points,
         `basketball_hoop`=basketball_points,
         `splashground_&_splashpad`=splashground_points) %>%
  pivot_longer(!year) %>%
  filter(year>=2018) %>%
  mutate(name2=str_remove(name,"points")) %>%
  mutate(name2=str_replace_all(name2,"_"," ")) %>%
  mutate(name2=str_to_sentence(name2))  -> p1data
```


```{r, fig.height=3.75, fig.width=3.75}
p1data %>%
  ggplot(aes(x=name2, y=value)) + 
  geom_point(aes(x=reorder(name2, value, median), y=value, color=factor(year)),position=position_jitter(0.15),size=1.5, alpha=0.5,shape=as.numeric(16)) +
  geom_boxplot(position=position_nudge(x=0.25, y=0),
               width=0.2, outlier.shape = NA, varwidth=FALSE, color="grey30", alpha=0.3)+
  coord_flip() + 
  theme_minimal(base_size=10) + 
  scale_color_manual(values=c("#ffcb69","#66c4b9","#392f5a")) + 
  scale_y_continuous(limits=c(-3,103),expand = c(0,0)) + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 13))+
  theme(axis.title = element_text(size=9, face="bold"),
        plot.margin = margin(15, 15, 10, 15),
        panel.grid.minor=element_blank(),
        panel.grid.major.y=element_blank(),
        legend.position="none",
        plot.title.position = "plot",
        plot.title=element_text(face="bold",size=24),
        plot.subtitle = element_markdown(size=13.1)) + 
  labs(y="Points",x="Measure",
       caption="\nTidyTuesday 26/2021 | Data from The Trust for Public Land",
       title="Public Parks in U.S. Cities",
       subtitle="Normalised values (points) of <span style = 'color:#ffcb69'><b>2018</b></span>, <span style = 'color:#66c4b9'><b>2019</b></span> and <span style = 'color:#392f5a'><b>2020</b></span><br>") 
  
```
ALT text: Dotted box plot showing the yearly normalized values (points) of public parks features in U.S cities for ten measures, where 2019 and 2020 have better points than 2018 across the measures. 


#### Ranking from 2012 to 2020 by city

```{r, fig.height=3, fig.width=3.75}
# bump chart
# reference: https://twitter.com/kierisi/status/1407171923633651717
parks2 = parks %>%
  mutate(park_pct_city_data=parse_number(park_pct_city_data),
         pct_near_park_data=parse_number(pct_near_park_data),
         spend_per_resident_data=parse_number(spend_per_resident_data),
         mean_spend_per_resident_data = mean(spend_per_resident_data),
         mean_diff= spend_per_resident_data-mean_spend_per_resident_data) %>%
  select(-city_dup)

# 7 most populous cities in United States
cities = c("New York","Los Angeles","Chicago","Houston","Phoenix",
           "Philadelphia","San Antonio")
parks2a= parks2 %>% filter(city %in% cities)

parks2a %>% 
  ggplot(aes(x=year,y=rank,color=city)) + 
  geom_point(size=2) + 
  geom_bump() +
  geom_text(data=parks3%>% filter(year==2020), aes(x=2020.2, y=rank, label=city, color=city),size=3, hjust=0) + 
  geom_text(data=parks3%>% filter(year==2012), aes(x=2010.8, y=rank, label=city, color=city),size=3, hjust=0) + 
  scale_y_reverse(breaks=c(1,21,41,61,81)) + 
  scale_x_continuous(limits=c(2010.8,2021),breaks=c(2012,2016,2020), position="top") +
  theme_minimal() + 
  theme(panel.grid.minor=element_blank(),
        legend.position="none") + 
  scale_color_d3() + 
  labs(x="",y="Rank")
```

```{r}
parks2c = parks2b %>% filter(year==2020) %>% arrange(-spend_per_resident_data) %>%
  mutate(city=fct_inorder(city))
levels(parks2c$city)
```

```{r}
parks2c %>% select(city,spend_per_resident_data)
```

#### Park spending by city

```{r}
# bump sankey hybrid using {ggsankey}
# reference:https://twitter.com/CedScherer/status/1384420013533241351/photo/1
# df with 10 most populous cities 
city10 = c("New York","Los Angeles","Chicago","Houston","Phoenix",
           "Philadelphia","San Antonio","San Diego","Dallas","San Jose")
parks2b= parks2 %>% filter(city %in% city10)

# colors
cols <- c("#c1121f", "#3C5488FF", "#e85d04","#e09f3e","#003b4c",
          "#778e99","#b2967d", "grey78","grey71","#343a40") 

# labels
df_labs <- tibble(
    city = factor(levels(parks2c$city), levels = levels(parks2c$city)),
    spend_per_resident_data =  c(600, 400, 200, 50, -100, 
                  -240, -360, -475, -580,-680)
  )

# plot
parks2b %>%
  ggplot(aes(x=year, value=spend_per_resident_data, node=city, fill=factor(city,levels=parks2c$city))) + 
  geom_sankey_bump(space=15, smooth=12) + 
  scale_fill_manual(values=cols) + 
  geom_text(data=df_labs, aes(x=2020.1, y=spend_per_resident_data, label=city),
            hjust=0,size=3,lineheight = .75, color=after_scale(cols), fontface="bold") +
  scale_x_continuous(limits=c(2012,2021)) +
  theme_void() + 
  theme(panel.grid.major.x = element_line(color = "#E8E1DB"),
        legend.position="none",
        axis.text.x=element_text(color="grey50",size=8)) 
```

#### Mean difference in park spending 
* Chicago, Houston, Los Angeles and New York

```{r}
# bar plot (mean difference)
# reference: https://twitter.com/jrosecalabrese/status/1407373878133936128/photo/1
parks2 %>% filter(city %in% c("New York","Los Angeles","Chicago","Houston")) %>%
  ggplot(aes(x=factor(year), y=mean_diff, group=city, fill=city)) + 
  geom_col(position="dodge") + 
  scale_fill_manual(values=c("#f6bd60","#f5cac3","#84a59d","#f28482")) +
  theme_classic(base_size=10) + 
  theme(panel.grid=element_blank(),
        legend.position="top")
```

#### Points by amenity type and city 
* Cities: Chicago, Houston, Los Angeles and New York
* Amenities: Basketball, Splashpads, Restrooms, Rec/Sr Centers, Playgrounds, Dog Parks

```{r}
# radar plot version 1
# reference: https://twitter.com/Juanma_MN/status/1407386167176220682/photo/1
dd= parks %>%
  filter(year == 2020) %>% 
  filter(city %in% c("New York","Los Angeles","Chicago","Houston")) %>%
  select(c('city', features[5:10])) %>% as.data.frame()

rownames(dd)<-dd$city #set facet title
data$city <- NULL
colors= c("#f6bd60","#f5cac3","#84a59d","#f28482")

par(mar=c(1, 2, 2, 1)) #decrease default margin
layout(matrix(1:4, ncol=2)) #draw 4 plots to device
#loop 
lapply(1:4, function(i) { 
    radarchart(rbind(rep(100,6), rep(0,6), dd[i,-1]),pty=32,
               pcol = '#89b0ae', pfcol = '#89b0ae',
             cglwd = .12, plwd = .001, title = row.names(dd[i,]), vlcex=.8,
             vlabels = c('Basketball', 'Splashpads', 'Restrooms', 'Rec/Sr Centers', 'Playgrounds', 'Dog Parks'))
})

```
```{r}
dd %>% rename_at(vars(matches("_points")), ~ str_remove(., "_points"))
```


```{r}
# radar plot v2
# function reference: https://www.datanovia.com/en/blog/beautiful-radar-chart-in-r-using-fmsb-and-ggplot-packages/

dd= dd %>% rename_at(vars(matches("_points")), ~ str_remove(., "_points")) 

create_beautiful_radarchart <- function(data, color = "#00AFBB", 
                                        vlabels = colnames(data), vlcex = 0.8,
                                        caxislabels = NULL, title = NULL, ...){
  radarchart(
    data, axistype = 1,
    # Customize the polygon
    pcol = color, pfcol = scales::alpha(color, 0.7), plwd = .001, plty = 1, pty=32,
    # Customize the grid
    cglcol = "black",  cglwd = .2,
    # Customize the axis
    axislabcol = "white", 
    # Variable labels
    vlcex = vlcex, vlabels = vlabels,
    caxislabels = caxislabels, title = title, ...
  )
}

# Define colors and titles
colors <- c("#E64B35FF","#4DBBD5FF","#00A087FF","#3C5488FF")
titles <- c("Chicago","New York","Los Angeles","Houston")

# Reduce plot margin using par()
# Split the screen in 3 parts
op <- par(mar = c(1, 1, 1, 1))
par(mfrow = c(2,2))

# Create the radar chart
lapply(1:4, function(i) { 
    create_beautiful_radarchart(
      data= rbind(rep(100,6), rep(0,6), dd[i,-1]),
      color = colors[i], title=titles[i], 
      )
})
par(op)
```

### Resident park spending in 2019 and 2020

```{r}
# resident park spending in 2019 and 2020
# prepare data
spend1= parks %>% 
  filter(year>=2019) %>%
  mutate(spend_per_resident_data=parse_number(spend_per_resident_data)) %>%
  select(year, city, spend_per_resident_data) %>%
  group_by(city) %>%
  filter(n() > 1) %>%
  arrange(city,year) %>%
  dplyr::mutate(diff = spend_per_resident_data- lag(spend_per_resident_data),
                col=ifelse(diff>0,"increase","decrease"),
                diff2 = abs(diff)) 
# get 10 cities with the most difference from 2019
spend_city = spend1 %>% arrange(-diff2) %>% ungroup() %>% slice(1:10)
spend2 = spend1 %>% filter(city %in% spend_city$city) %>% arrange(city,year)
# data for geom_text
spend3 = spend2 %>% group_by(city) %>% dplyr::mutate(mea = median(spend_per_resident_data)) %>%
  ungroup() %>% mutate(diff3=if_else(diff>0,paste0("+",diff), paste0(diff))) %>% filter(year==2020)
# city levels order by difference
s4 = spend3 %>% arrange(diff2) %>% mutate(city=fct_inorder(city)) 

# plot
spend2 %>% ggplot() + 
  geom_point(aes(x=spend_per_resident_data,y=factor(city,s4$city),
                 shape=factor(year)),size=2.5) + 
  geom_line(aes(group=city, x=spend_per_resident_data, y=city, color=col),show.legend = F) +
  scale_color_manual(values=c("#a4243b","grey"),na.value="#2b9348") + 
  theme_minimal(base_size=10) + 
  scale_shape_manual(values=c(1,19)) +
  geom_text(data=spend3,aes(x=mea, y=city, label=diff3),size=2.7, vjust=-0.7) + 
  theme(panel.grid.minor=element_blank(),
        plot.title.position = "plot") + 
  labs(x="Spend per resident (in USD)", y= "City", shape="Year",
       title="Park spending per resident in 2019 and 2020")
```



