---
title: "Tidy Tuesday Week 31/2021"
date: "2021/07/27"
output: html_notebook
---

Data visualization exercise

This notebook uses [#TidyTuesday](https://github.com/rfordatascience/tidytuesday) week 30 [Olympic Medals](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-07-27/readme.md), data from [Kaggle](https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results).

```{r}
# load libraries 
library(tidyverse)
library(ggdark)
library(patchwork)
library(scales)
library(glue)
library(maps)
library(mapdata)
library(countrycode)
library(geofacet)
library(colorspace)
library(ggsci)
library(ggdist)
library(gghalves)
library(tidyquant)
library(ggtext)
library(ggmosaic)
library(ggflags)
library(DT)
```

```{r}
# import data
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')
```

#### Cumulative gold medals, by NOC (country)
* shared on [Twitter](https://twitter.com/leeolney3/status/1419918423744782379)

```{r}
gold_cumulative = olympics %>% filter(medal=="Gold") %>% 
  group_by(noc, year) %>% tally() %>%
  mutate(cumulative = cumsum(n)) 
```

```{r}
summary(gold_cumulative$year)
```

```{r}
gold_cumulative %>% 
  group_by(noc) %>% 
  summarise(max_n = max(cumulative)) %>%
  arrange(desc(max_n))
```

```{r}
# main plot
mainplot = gold_cumulative %>% 
  mutate(color=case_when(noc=="USA" ~ "#e63946",
                         noc=="URS" ~ "grey70",
                         noc=="GER" ~ "#fb8500",
                         noc=="GBR" ~ "#4cc9f0",
                         noc=="ITA" ~ "#90be6d",
                         TRUE~"#495057")) %>%
  ggplot() +
  aes(x=year, y= cumulative, color=color) + 
  geom_path() + 
  scale_color_identity(guide="none") + 
  scale_x_continuous(limits=c(1896,2030), expand=c(0,0), breaks=seq(1900,2016,25)) +
  annotate("text", y=2638, x=2017, label="USA: 2638", hjust=0,size=3, color="#e63946") +
  annotate("text", y=1082, x=1989, label="URS: 1082", hjust=0,size=2.5, color="grey70") +
  annotate("text", y=745, x=2017, label="GER: 745", hjust=0,size=2.5, color="#fb8500") +
  annotate("text", y=666, x=2017, label="GBR: 678", hjust=0,size=2.5, color="#4cc9f0") +
  annotate("text", y=575, x=2017, label="ITA: 575", hjust=0,size=2.5, color="#90be6d") +
  geom_segment(aes(x = 1900, y = 0, xend = 1900, yend = 200), col = "grey70", linetype = "dotted", size=.2) +
  geom_segment(aes(x = 1925, y = 0, xend = 1925, yend = 650), col = "grey70", linetype = "dotted", size=.2) +
  geom_segment(aes(x = 1950, y = 0, xend = 1950, yend = 1000), col = "grey70", linetype = "dotted", size=.2) +
  geom_segment(aes(x = 1975, y = 0, xend = 1975, yend = 1600), col = "grey70", linetype = "dotted", size=.2) +
  geom_segment(aes(x = 2000, y = 0, xend = 2000, yend = 2300), col = "grey70", linetype = "dotted", size=.2) +
  theme_minimal() +
  theme(panel.grid=element_blank(),
        plot.margin=unit(c(0.5,0.5,0.5,0.5),"cm"),
        axis.title=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.text.x=element_text(margin=margin(t=-10), size=7, color="grey90"),
        plot.caption=element_text(size=8, color="grey80"),
        plot.background = element_rect(fill = "#212529", color=NA),
        plot.title=element_text(margin=margin(t=15,b=-30),hjust=0,face="bold",size=24,color="grey90"),
        plot.subtitle=element_text(margin=margin(t=35,b=-10),hjust=0,face="bold",size=8.5,color="grey80")
        ) +
  ggtitle("Olympic Gold Medals") + 
  labs(caption="\n#TidyTuesday Week 31 | Data from Kaggle",
       subtitle="Cumulative Olympic gold medals won, by National Olympic Committee (NOC) from 1896 to 2016")
```


```{r}
# map insert
world <- map_data("world") %>% filter(region != "Antarctica")
world2 = world %>% mutate(col = case_when(region=="UK"~"#4cc9f0",
                         region=="Germany"~"#fb8500",
                         region=="USA"~"#e63946",
                         region=="Italy"~"#90be6d",
                         TRUE~"#495057"
                         ))

map_insert = ggplot() + geom_map(data=world2, map=world2, aes(long, lat, map_id=region, fill=col)) + 
  scale_fill_identity() + 
  theme_void() + 
  theme(plot.background = element_rect(fill = "#212529", color=NA)) +
  coord_quickmap()
```


```{r}
# insert map 
mainplot |inset_element(map_insert ,
                    align_to = "full",
                    clip = FALSE,
                    on_top = TRUE,
                    ignore_tag = TRUE,
                    left = 0.01, 
                    bottom = 0.15, 
                    right = 0.5, 
                    top = 1) 
```
ALT text: Line plot showing the cumulative gold medals won by respective NOCs from 1896 to 2016, where USA won the most gold medals (n=2638). 

#### Ratio of medals to participation
```{r}
# medal (4 levels) ratio
medal4 = olympics %>% 
  mutate(country= countrycode(noc, origin='ioc', destination='country.name')) %>% # get country name 
  count(country, year, medal) %>%
  group_by(country, year) %>% mutate(sum_medal=sum(n)) %>%
  ungroup() %>%
  mutate(ratio = n/sum_medal)
```

```{r}
setdiff(medal2$region, world$region)
```

```{r}
# medal (2 levels) ratio
medal2 = olympics %>%
  mutate(region= countrycode(noc, origin='ioc', destination='country.name')) %>% # get country name
  mutate(medal2 = ifelse(is.na(medal),"0","1")) %>%
  group_by(region, noc, medal2) %>% tally() %>%
  mutate(ratio=n/sum(n)*100) %>%
  filter(medal2=="1") %>%
  drop_na() %>%
  mutate(ratio_cat = cut(ratio, breaks=seq(0,30,10))) %>%
  mutate(region= case_when(region=="United States"~"USA",
                            region=="United Kingdom"~"UK",
                            region=="Côte d’Ivoire"~"Ivory Coast",
                            region=="Czechia"~"Czech Republic",
                            TRUE ~ region)) 
medal2
```


```{r, warning=F}
joined = left_join(world, medal2, by="region") 

ggplot() + geom_map(data=joined, map=joined, aes(long, lat, map_id=region, fill=ratio_cat)) + 
  theme_void() + 
  coord_quickmap() + 
  theme(legend.position="top",
        plot.title=element_text(hjust=0.5, size=10, face="bold"),
        plot.margin=unit(c(0.5,1,0.5,1),"cm"),
        legend.title=element_text(size=8),
        legend.text=element_text(size = 8)) + 
  scale_fill_manual(values=c("#219ebc","#ffb703","#e63946"),na.value="grey",
                    breaks=c("(0,10]","(10,20]","(20,30]")) + # hide NA in legend
  guides(fill = guide_legend(title = "Ratio (%)",
                             title.position = "left",
                             title.hjust = 0.5,
                             label.position = "bottom",
                             label.hjust = 0.5,
                             nrow = 1,
                             keyheight = .5,
                             keywidth = 4)) + 
  labs(title="Ratio of medals to participation, by country from 1896 to 2016\n")
```




#### Distribution of medalist height by sport 

```{r}
sport3 = olympics %>% filter(!is.na(medal)) %>% count(sport, sort=T) %>% slice(1:5)

df_height = olympics %>% 
  filter(!is.na(medal)) %>%
  filter(sport %in% sport3$sport) %>%
  drop_na(height, sport) %>%
  mutate(sport=fct_infreq(sport)) %>%
  mutate(sport_num = as.numeric(sport))
  
df_height_stat = df_height %>%
  group_by(sport, sport_num) %>%
  summarise(median=median(height),
            max=max(height),
            n=n())
```


```{r}
# raincloud v1: stat_half_eye and geom_half_point
# reference: https://z3tt.github.io/OutlierConf2021/
df_height %>%
  ggplot(aes(x=sport_num, y=height, color=sport)) + 
  geom_half_point(side="l", size=0.7, shape=21, alpha=0.4, range_scale = 0.3) + 
  stat_summary(
    geom = "linerange",
    fun.min = function(x) -Inf,
    fun.max = function(x) median(x, na.rm = TRUE),
    linetype = "dotted",
    orientation = "x",
    size = .5
  ) +
  ggdist::stat_halfeye(
    aes(x = sport_num,color = sport,fill = after_scale(colorspace::lighten(color, .3))
    ),shape = 18,point_size = 3,interval_size = 1.8,adjust = .5,.width = c(0, 1)) + 
  geom_text(data= df_height_stat,
    aes(y = median, x=sport_num,label = median),
    color = "white",fontface = "bold",size = 3,nudge_x = .15) +
  geom_text(data= df_height_stat,
    aes(y = max, x=sport_num,label = glue::glue("n = {n}")),
    fontface = "bold",size = 3,nudge_x = .15, hjust=0) +
  scale_fill_futurama() +
  scale_color_futurama() +
  scale_x_continuous(breaks = 1:5, 
                     labels = c("Athletics","Swimming","Rowing","Gymnastics","Fencing")) +
  scale_y_continuous(expand=c(.1,.1), breaks=seq(140,210,10)) +
  coord_flip() + 
  theme(legend.position="none",
        panel.grid.minor=element_blank(),
        panel.grid.major.y=element_blank(),
        plot.margin=unit(c(0.5,2,0.5,0.5),"cm"),
        plot.title.position = "plot",
        axis.title.y=element_blank(),
        axis.title.x=element_text(size=9),
        panel.grid.major.x=element_line(size=.3),
        axis.text.y=element_text(size=9)) +
  labs(subtitle="Distribution of Olympics medalist height, by sport (5 sports with highest medal count)",
       y="Height (in cm)")
```


```{r, warning=F, fig.height=3.8}
# raincloud v2: stat_half_eye, geom_boxplot and stat_dots
# reference: https://www.r-bloggers.com/2021/07/ggdist-make-a-raincloud-plot-to-visualize-distribution-in-ggplot2/
df_height %>% 
  ggplot(aes(x=sport, y=height, fill=sport, color=sport)) + 
  stat_halfeye(adjust=0.5, justification=-.2, .width=0, point_color=NA) + 
  geom_boxplot(width=0.1, outlier.color=NA, alpha=0.5) + 
  stat_dots(side="left", justification=1.1, binwidth=unit(c(0, 0.01), "npc")) + 
  #gghalves::geom_half_point(side="l", alpha=.3, shape="|", size=5) + #range_scale=.4, 
  geom_text(data=df_height_stat, aes(y=median, x=sport, label=median), 
            fontface="bold",color="white",nudge_x = .3,size=4) +
  geom_text(data=df_height_stat, aes(y=max, x=sport, color=sport, label=glue::glue("n = {n}")), 
            fontface="bold",nudge_x = .3,size=4, nudge_y=3) +
  theme_tq(base_size=14) +
  theme(legend.position = "none",
        plot.margin=unit(c(0.5,2,0.5,0.5),"cm")) + 
  scale_fill_futurama() + 
  scale_color_futurama() +
  coord_flip() + 
  labs(title="Distribution of Olympics medalist height by sport",
       subtitle="Top 5 sports by medal count",
       x="Sport", y="Height (in cm)")
  
```

#### Swimming medals 
```{r}
olympics %>% 
  filter(!is.na(medal)) %>%
  filter(sport =="Swimming") %>%
  mutate(event_dist = parse_number(event)) %>%
  mutate(event=str_to_lower(event)) %>%
  filter(!str_detect(event, 'yard')) %>%
  drop_na() %>%
  mutate(type=case_when(str_detect(event, "relay")~"relay",
                        TRUE ~ "individual")) %>%
  mutate(noc=fct_lump(noc, 3)) %>%
  group_by(type, noc) %>% tally() %>% mutate(prop=round(n/sum(n),3)) %>%
  arrange(type, desc(n))
```

#### Female and male athletes
```{r}
mdata = olympics %>% 
  group_by(year, sex) %>%
  tally() %>% 
  mutate(prop=n/sum(n)) 

ggplot(olympics) +
  geom_mosaic(aes(x=product(sex, year), fill=sex), show.legend=F) + 
  scale_y_continuous(position="right") +
  theme(axis.text.x=element_text(angle=90, vjust=0.5,size=7, color="black"),
        axis.text.y=element_text(color="black",size=7),
        panel.grid.minor=element_blank(),
        axis.title=element_blank(),
        axis.ticks=element_line(color="grey"),
        axis.ticks.length = unit(.4, "cm"),
        panel.grid.major.x=element_blank(),
        plot.margin=unit(c(1,1,1,1),"cm"),
        plot.title=element_markdown(face="bold",size=12),
        plot.subtitle = element_text(size=8, color="grey30")) + 
  coord_cartesian(expand = FALSE,clip="off") + 
  scale_fill_manual(values=c("#ee9b00","#656d4a")) + 
  labs(title="Olympics <span style = 'color:#ee9b00'>female</span> and <span style = 'color:#656d4a'>male</span> athletes, by year",
       subtitle="Cross sectional proportion of athletes by gender, from 1896 to 2016. Winter and Summer Games were held in the same year up until\n1992. After that, they staggered them such that Winter Games occur on a four year cycle starting with 1994, then Summer in 1996,\nthenWinter in 1998, and so on.\n")

```

#### Event count by season and year

```{r}
olympics %>%
  count(season, year, event) %>%
  group_by(season, year) %>% tally(n) %>%
  ggplot(aes(x=year, y=n, fill=season)) + 
  geom_col() + 
  scale_x_continuous(breaks=seq(1896,2016,20)) + 
  scale_y_continuous(labels=scales::comma) + 
  theme(panel.grid.minor=element_blank(),
        legend.position = "top",
        plot.margin=unit(c(1,2,1,1),"cm"),
        axis.title=element_text(size=9),
        plot.title.position="plot",
        legend.justification = "left",
        legend.margin = margin(l=-45),
        legend.key.width =unit(.25,"cm"),
        panel.grid.major=element_line(size=.3)) + 
  scale_fill_npg() + 
  labs(fill="", x="Year", y="Event count",
       subtitle="Olympic event count, by season and year") 
```


#### 10 countries with most summer olympics medal

```{r}
s10 = olympics %>% 
  filter(season=="Summer",year>1990,!is.na(medal)) %>% 
  count(noc, sort=T) %>%
  slice(1:10)
```

```{r}
# reference: https://kpress.dev/blog/2021-07-26-tidy-tuesday-olympic-medals/
# reference: https://github.com/davidsjoberg/ggbump
df_rank = olympics %>% 
  filter(season=="Summer",year>1990,!is.na(medal)) %>%
  filter(noc %in% s10$noc) %>%
  mutate(region= countrycode(noc, origin='ioc', destination='genc2c')) %>% 
  mutate(region= str_to_lower(region)) %>%
  group_by(year, region) %>% tally() %>%
  arrange(year, -n) %>%
  mutate(rank = dense_rank(desc(n)))

df_rank %>%
  ggplot(aes(year, rank, color=region)) + 
  geom_point(size=2) +
  geom_bump(size=1) + 
  geom_flag(data= df_rank %>% filter(year==2016),aes(country=region),size=8) +
  #geom_text(data= df_rank %>% filter(year==2016), aes(x=year+0.5, y=rank, label=country, hjust=0),size=3) +
  scale_y_reverse(breaks=seq(1,10,1)) + 
  scale_x_continuous(breaks=seq(1992, 2016,4)) + 
  scale_color_d3() +
  theme(panel.grid.minor=element_blank(),
        panel.grid.major=element_blank(),
        plot.margin=unit(c(.5,1,.5,.5),"cm"),
        legend.position="none",
        plot.background=element_rect(fill="#e9ecef", color="NA"),
        axis.title=element_text(size=8.5),
        plot.title.position = "plot",
        axis.text=element_text(color="black")) + 
  labs(x="Year", y="Rank",
       subtitle="Countries with most summer Olympics medals\n") 
```


#### Sport: min year, max year, event count
  
```{r}
olympics %>% group_by(sport) %>%
  summarise(min_year= min(year), max_year=max(year), year_count = n_distinct(year)) %>%
  arrange(min_year, max_year) %>%
  ungroup() %>%
  DT::datatable(rownames=FALSE,options = list(order = list(list(1, 'asc'))))
```


