---
title: "Tidy Tuesday Week 36/2021"
date: "2021/08/31"
output: html_notebook
---

[Tidy Tuesday](https://github.com/rfordatascience/tidytuesday) Week 36 [Bird Baths](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-08-31/readme.md), Data from [Cleary et al, 2016](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0150899#abstract0).

```{r}
# Load libraries
library(tidyverse)
library(ggsci)
library(hrbrthemes)
library(ggtext)
library(waffle)
library(tidymodels)
library(patchwork)

options(dplyr.summarise.inform = FALSE)
```


```{r}
# Import data
bird_baths <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-31/bird_baths.csv',show_col_types = FALSE)
```

```{r}
# Missing data
colSums(is.na(bird_baths))
# Variable unique value count 
bird_baths %>% summarise_all(n_distinct)
```

### Bird count by region and location  
* shared on [Twitter](https://twitter.com/leeolney3/status/1432635306134720514)   
```{r, fig.width=3.5, fig.height=2.5}
bird_baths %>% drop_na(survey_year) %>%
  group_by(bioregions, urban_rural, survey_year) %>%
  tally(bird_count) %>%
  ungroup() %>%
  mutate(survey_year2=ifelse(survey_year==2015, "Summer'15", "Winter'14")) %>%
  mutate(group=paste(survey_year2,urban_rural)) %>%
  mutate(group2=factor(group, levels=c("Winter'14 Rural","Winter'14 Urban","Summer'15 Rural", "Summer'15 Urban"), ordered=T)) %>%
  ggplot(aes(y=reorder(bioregions,n), x=n, color=group2, shape=group2)) +
  geom_line(aes(group=bioregions), color="grey70") +
  geom_point(size=2.2) +
  scale_shape_manual(values=c(16,16,17,17)) +
  scale_x_continuous(limits=c(0,1200), expand=c(0,0)) +
  scale_color_futurama() +
  theme_ipsum(base_size=9) +
  theme(legend.position = "top",
        panel.grid.major.y = element_blank(),
        legend.text = element_text(size=9),
        legend.title = element_text(size=9),
        legend.justification = "left",
        legend.box.margin = margin(l=-111.5, b=-5),
        axis.title.x=element_text(hjust=0.5, size=9),
        axis.title.y=element_text(hjust=0.5, size=9),
        plot.title.position = "plot",
        plot.title=element_text(size=14),
        plot.subtitle = element_text(size=9, margin=margin(t=-4), color="grey20"),
        plot.margin=unit(c(.5,1.5,.5,.5),"cm"),
        plot.caption=element_text(size=8, color="grey20", face="plain")
        ) + 
  labs(x= "Bird count", 
       y= "Region",
       color="Period and location", shape="Period and location",
       title="Bird occurrence count at bird baths, in Australia",
       caption="Week 36 Bird Baths | Data from Cleary et al, 2016",
       subtitle="Data on bird occurrence at bird baths were collected during austral winter (June 24th to July 26th 2014) and summer (January 27th to\nFebruary 29th 2015). The highest bird count (n=1190) at bird baths was recorded during Winter'14 in Sydney Basin rural location, and\nlowest bird count (n=26) was recorded during Summer'15 in Victorian Midlands urban location.")
```
ALT text: The highest bird count (n=1190) at bird baths was recorded during Winter'14 in Sydney Basin rural location and lowest bird count (n=26) was recorded during Summer'15 in Victorian Midlands Urban location. 

```{r}
# bird count by year
bird_baths  %>% group_by(survey_year) %>% tally(bird_count)

# bird count by region
bird_baths %>% 
  drop_na(bioregions) %>%
  group_by(bioregions) %>%
  tally(bird_count, sort=T)

# bird count by species
bird_baths %>%
  drop_na(bird_type) %>%
  group_by(bird_type) %>% 
  tally(bird_count, sort=T)
```

### Species count by region and period
```{r}
bird1 = bird_baths %>% 
  drop_na() %>%
  group_by(bioregions,survey_year, bird_type) %>%
  tally(bird_count) %>%
  filter(n>0) %>%
  ungroup() %>%
  group_by(bioregions,survey_year) %>%
  summarise(n_type = n_distinct(bird_type)) %>%
  ungroup() 

bird1 %>%
  mutate(n=ifelse(survey_year==2014,-1*n_type,n_type)) %>%
  ggplot(aes(y=bioregions, x=n, fill=factor(survey_year))) + 
  geom_col(width=.65) +
  geom_text(aes(label=abs(n), hjust=ifelse(n>0,1.3,-0.5)), 
            fontface="bold", color="white", size=3.1) +
  #scale_x_continuous(limits=c(-110,110),expand=c(0,0)) +
  scale_x_continuous(limits=c(-110,110), expand=c(0,0), labels=abs) +
  scale_fill_manual(values=c("#404e7c","#71b48d")) +
  theme_minimal(base_size=10) + 
  theme(legend.position = "none",
        panel.grid=element_line(linetype = "dotted",color="grey50", size=.2),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank(),
        axis.title=element_blank(),
        axis.text=element_text(size=9.1),
        plot.margin=unit(c(1,1.5,1,1),"cm"),
        plot.title.position = "plot",
        plot.title=element_text(face="bold", size=12),
        plot.subtitle=element_markdown(face="bold",size=11, hjust=0.5,
                                       margin = margin(10,0,5,129))
        ) + 
  labs(subtitle="<span style='color:#404e7c;'>Winter 2014</span> | <span style='color:#71b48d;'>Summer 2015</span>",
       title="Bird baths in Australia: Species count by bioregion and period")
```

### Species count by bioregion (2014 and 2015)   
* code reference: https://twitter.com/nicci_potts/status/1432797343821312006   
* Australia bioregion shapefiles source: http://www.environment.gov.au/fed/catalog/search/resource/downloadData.page?uuid=%7B2CD8B271-A723-4B1B-8D26-9C68B1974B45%7D   

```{r}
library(tmap)
library(rgdal)
library(sf)
library(rmapshaper)
library(cowplot)
```

```{r}
# NSW
NSWLayer <- st_read("ibra61_reg_nsw_shape/ibra61_reg_nsw_shape.shp", layer= "ibra61_reg_nsw_shape") 
NSWLayer <- st_transform(NSWLayer, "+init=epsg:4326") 
SimpleNSW <- ms_simplify(NSWLayer) 

# Victoria
VICLayer <- st_read("ibra61_reg_vic_shape/ibra61_reg_vic_shape.shp", layer= "ibra61_reg_vic_shape") 
VICLayer <- st_transform(VICLayer, "+init=epsg:4326") 
SimpleVIC <- ms_simplify(VICLayer)

# South Australia
SALayer <- st_read("ibra61_reg_sa_shape/ibra61_reg_sa_shape.shp", layer= "ibra61_reg_sa_shape") 
SALayer <- st_transform(SALayer, "+init=epsg:4326") 
SimpleSA <- ms_simplify(SALayer) 

# Queensland
QLDLayer <- st_read("ibra61_reg_qld_shape/ibra61_reg_qld_shape.shp", layer= "ibra61_reg_qld_shape") 
QLDLayer <- st_transform(QLDLayer, "+init=epsg:4326") 
SimpleQLD <- ms_simplify(QLDLayer)

# join shapefile
bio <- rbind(SimpleNSW, SimpleVIC, SimpleSA, SimpleQLD)
```

```{r}
# prepare data
bird_df = bird1 %>% 
  rename("REG_NAME" = bioregions) %>%
  group_by(REG_NAME) %>% tally(n_type)
# merge
bird_map = merge(bio,bird_df, by= 'REG_NAME')
```

```{r}
# plot map
#map
map <- tm_shape(bird_map) +
  tm_polygons(
    col = "n",
    palette = 'Greens',
    title = '',
    lwd = .25,
    border.col = 'gray20'
  ) +
  tm_text("REG_CODE",
          size= "AREA")+
  tm_style("white") +
  tm_layout(
    frame = FALSE,
    bg.color = '#3d5a80',
    "Bird baths in Australia:\nSpecies count\nby bioregion",
    title.color = 'white',
    title.size = 1,
    legend.bg.color = TRUE,
    legend.frame = FALSE,
    legend.position = c("left", "bottom"),
    legend.text.color = 'white',
    legend.text.size = 0.7
  ) 
map
```

### Top species by region and period
```{r}
bird_baths %>% 
  drop_na() %>%
  group_by(bioregions,survey_year, bird_type) %>%
  tally(bird_count) %>%
  filter(n>0) %>%
  ungroup() %>%
  group_by(bioregions, survey_year) %>%
  mutate(perc=paste0(sprintf("%4.1f", n / sum(n) * 100), "%")) %>%
  top_n(1,perc)
```

### Bird count by bioregion and location
```{r}
count = bird_baths %>% 
  drop_na() %>%
  group_by(bioregions,urban_rural) %>%
  tally(bird_count) %>%
  mutate(n=n/10)

count_order = count %>% 
  group_by(bioregions) %>% tally(n) %>%
  arrange(desc(n)) %>%
  mutate(bioregions=fct_inorder(bioregions))
```

```{r}
count %>% mutate(bioregions=factor(bioregions, levels=count_order$bioregions)) %>%
  ggplot(aes(fill=urban_rural, values=n)) + 
  geom_waffle(color="white", size=.25, n_rows=5, flip=T, show.legend=T) +
  facet_wrap(~bioregions, nrow=1, strip.position = "bottom",
             labeller = labeller(bioregions = label_wrap_gen(width = 9))) + 
  scale_x_discrete(expand=c(.25,.25)) + 
  scale_y_continuous(labels = function(x) x * 50, # multiplyer same as n_rows
                     expand = c(0,0)) +
  scale_fill_manual(values=c("#588b8b","#f28f3b")) +
  coord_equal() +
  guides(fill = guide_legend(reverse=T,keywidth = unit(.5, "cm"), keyheight = unit(.5, "cm"))) +
  theme_minimal(base_size = 8.5) + 
  theme(panel.grid=element_line(size=.3,color="grey80"),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank(),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        legend.position = "top",
        panel.spacing.x = unit(0, "lines"),
        axis.text.y=element_text(margin=margin(r=-20), vjust=-0.7, color="grey30"),
        strip.text=element_text(color="grey30", size=7.5),
        axis.ticks.y=element_line(color="grey80",size=.3),
        axis.ticks.length.y.left = unit(1, "cm"),
        legend.title=element_blank(),
        legend.justification = "left"
        ) + 
  labs(title="Bird Baths in Australia: Count of bird occurances by bioregion and location",
       subtitle="Count of bird occurances by bioregion and location, where each square represents 10 occurances observed.")
```
### Probablity and predicted probabilty of seeing bird type   
* reference: https://juliasilge.com/blog/bird-baths/

```{r}
# theme
theme = theme_minimal(base_size=10) +
  theme(panel.grid=element_line(size=.3),
        plot.title=element_text(size=10),
        plot.title.position = "plot",
        legend.position = "bottom",
        panel.grid.minor = element_blank(),
        legend.box.margin= margin(t=-10))
```


```{r}
# top 15 bird species by count
top_birds <-
  bird_baths %>%
  filter(is.na(urban_rural)) %>%
  arrange(-bird_count) %>%
  slice_max(bird_count, n = 15) %>%
  pull(bird_type)

# summary 
bird_parsed <-
  bird_baths %>%
  filter(
    !is.na(urban_rural),
    bird_type %in% top_birds
  ) %>%
  group_by(urban_rural, bird_type) %>%
  summarise(bird_count = mean(bird_count), .groups = "drop")
```

```{r}
# plot probability
p1 <-
  bird_parsed %>%
  ggplot(aes(bird_count, bird_type)) +
  geom_segment(
    data = bird_parsed %>%
      pivot_wider(
        names_from = urban_rural,
        values_from = bird_count
      ),
    aes(x = Rural, xend = Urban, y = bird_type, yend = bird_type),
    alpha = 0.7, color = "gray70", size = 1
  ) +
  geom_point(aes(color = urban_rural), size = 2) +
  scale_x_continuous(labels = scales::percent, limits=c(0,0.45), expand=c(0,0)) +
  scale_color_npg(guide = guide_legend(reverse = TRUE)) +
  theme + 
  labs(x = NULL, y = NULL, color = NULL,shape=NULL,
       title="Probability of seeing bird")
```

```{r}
# prepare data for model 
bird_df <-
  bird_baths %>%
  filter(
    !is.na(urban_rural),
    bird_type %in% top_birds
  ) %>%
  mutate(bird_count = if_else(bird_count > 0, "bird", "no bird")) %>%
  mutate_if(is.character, as.factor)
```

```{r}
# 10-fold cv
set.seed(123)
bird_split <- initial_split(bird_df, strata = bird_count)
bird_train <- training(bird_split)
bird_test <- testing(bird_split)

set.seed(234)
bird_folds <- vfold_cv(bird_train, strata = bird_count)
```

```{r}
# glm
glm_spec <- logistic_reg()

# preprocess
rec_basic <-
  recipe(bird_count ~ urban_rural + bird_type, data = bird_train) %>%
  step_dummy(all_nominal_predictors())

# workflow
wf_basic <- workflow(rec_basic, glm_spec)
```

```{r}
# fit to train (10 times to 10 fold) 
doParallel::registerDoParallel()
ctrl_preds <- control_resamples(save_pred = TRUE)
rs_basic <- fit_resamples(wf_basic, bird_folds, control = ctrl_preds)

collect_metrics(rs_basic)
#augment(rs_basic) %>% roc_curve(bird_count, .pred_bird) %>% autoplot()
```
```{r}
# with explicit interaction terms
rec_interact <-
  rec_basic %>%
  step_interact(~ starts_with("urban_rural"):starts_with("bird_type"))

wf_interact <- workflow(rec_interact, glm_spec)
rs_interact <- fit_resamples(wf_interact, bird_folds, control = ctrl_preds)

collect_metrics(rs_interact)
#augment(rs_interact) %>% roc_curve(bird_count, .pred_bird) %>% autoplot()
```

```{r}
# fit to entire training set
bird_fit <- fit(wf_interact, bird_train)

new_bird_data <-
  tibble(bird_type = top_birds) %>%
  crossing(urban_rural = c("Urban", "Rural"))

# predict 
bird_preds <-
  augment(bird_fit, new_bird_data) %>%
  bind_cols(
    predict(bird_fit, new_bird_data, type = "conf_int")
  )
```

```{r}
# plot predicted probabilty 
p2 <-
  bird_preds %>%
  ggplot(aes(.pred_bird, bird_type, color = urban_rural)) +
  geom_errorbar(aes(
    xmin = .pred_lower_bird,
    xmax = .pred_upper_bird
  ),
  width = .15, size = 1.2, alpha = 0.5
  ) +
  geom_point(size = 2) +
  theme+
  scale_x_continuous(labels = scales::percent) +
  scale_color_npg(guide = guide_legend(reverse = TRUE)) +
  labs(title = "Predicted probability of seeing bird", y = NULL, color = NULL, x=NULL) 
```

```{r}
# combine plots
p1 + p2
```

### Bird sightings and species count by bioregion
* reference: https://twitter.com/geokaramanis/status/1433785579074998279/photo/1

```{r}
library(tidyverse)
library(here)
library(sf)
library(ggrepel)
library(biscale)
library(cowplot)
```

```{r}
# shp file: https://www.environment.gov.au/fed/catalog/search/resource/details.page?uuid=%7B1273FBE2-F266-4F3F-895D-C1E45D77CAF5%7D 
ibra <- read_sf(here::here("IBRA7_subregions_states", "IBRA7_subregions_states.shp")) %>% 
  st_make_valid() %>% 
  st_simplify(dTolerance = 0.1)
```

```{r}
# summary
summary = bird_baths %>% 
  filter(!is.na(bioregions)) %>% 
  group_by(bioregions) %>%
  filter(bird_count>0) %>%
  summarise(n=sum(bird_count), n_type=n_distinct(bird_type)) %>%
  mutate(bioregions = if_else(bioregions == "Victorian Volcanic Plain", "Southern Volcanic Plain", bioregions)) %>%
  bi_class(x = n, y = n_type , style = "quantile", dim = 3)

ibra_sf <- ibra %>% 
  filter(REG_NAME_7 %in% summary$bioregions) %>% 
  group_by(REG_NAME_7) %>% 
  summarise(geometry = st_union(geometry)) %>% 
  ungroup() %>% 
  rename(bioregions = REG_NAME_7)

join_sf = ibra_sf %>% left_join(summary)
```

```{r}
# centroid coordinates
totals = bird_baths %>% 
  filter(!is.na(bioregions)) %>% 
  group_by(bioregions) %>% 
  summarise(n = sum(bird_count)) 

ibra_centroids <- ibra_sf %>%
  st_centroid() %>% 
  left_join(totals) %>% 
  cbind(st_coordinates(.))
```


```{r}
p3 = ggplot() +
  geom_sf(data = ibra_sf, fill = "grey72", color = "grey99", size = 0.25) + 
  geom_sf(data = join_sf, aes(fill=bi_class), color = "grey99", size = 0.25) + 
  bi_scale_fill(pal = "GrPink", dim = 3, guide = "none") +
  geom_label_repel(data = ibra_centroids, aes(x = X, y = Y, label = bioregions), color = "grey20", point.padding = 10, size = 2, fontface = "bold", label.size = 0, fill = "#f8f8ff70", stat = "unique", nudge_y = 0) +
  theme_void() +
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"))
```

```{r}
p4 = bi_legend(pal = "GrPink", #DkCyan
            dim = 3,
            xlab = "Bird count",
            ylab = "Species count ",
            size = 2.5) + 
  theme(panel.border = element_blank(),
        axis.text = element_blank(),
        axis.title.x = element_text(size = 5,
                                    color = "#22222b", margin=margin(t=-5)),
        axis.title.y = element_text(size = 5,
                                    color = "#22222b", margin=margin(r=-5)),
        legend.text = element_text(size = 5),
        legend.text.align = 0)
```

```{r}
ggdraw() +
  draw_plot(p3, 0, 0, 1, 1) +
  draw_plot(p4, 0, 0.1, 0.25, 0.25) +
  draw_label("Bird sightings and species count at bird baths, in Australia", #fontface= "bold",
             color = "#22222b", size = 9, angle = 0, x =0.06, y = 0.95, hjust=0) 
```

### Bird sightings count by bioregion 
* reference: https://twitter.com/geokaramanis/status/1433785579074998279/photo/1
```{r}
summary(ibra_centroids$n)
```

```{r}
ggplot() +
  geom_sf(data = ibra_sf, fill = "grey72", color = "grey99", size = 0.25) +
  coord_sf(clip = "off", xlim = c(133, 155), ylim = c(-22, -40), expand = FALSE) + 
  geom_point(data = ibra_centroids, aes(X, Y, size=n, color=n)) + 
  geom_label_repel(data = ibra_centroids, aes(x = if_else(str_detect(bioregions, "Coastal"), X+.5, X - .5), y = Y, label = bioregions, hjust = if_else(str_detect(bioregions, "Coastal"), 0, 1)), color = "grey20", point.padding = 10, size = 2, fontface = "bold", label.size = 0, fill = "#f8f8ff70", stat = "unique", nudge_y = 0.15) +
  scale_size_continuous(limits=c(101, 1962), breaks=c(101, 500, 1000, 1500, 1962),
                        range = c(2,7)) +
  scale_color_continuous_sequential(limits=c(101, 1962),breaks=c(101, 500, 1000, 1500, 1962),
                                    palette = "viridis", rev=F) + 
  guides(color=guide_legend(), size=guide_legend()) +
  theme_void(base_size = 9) + 
  labs(color="", size="",
       title="Bird sightings at bird baths in Australia, by bioregions")
```



