---
title: "2023_05"
format: html
editor: visual
---

[TidyTuesday](https://github.com/rfordatascience/tidytuesday) week 5 [Pet Cats UK](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-31/readme.md), data from the Movebank for [Movebank for Animal Tracking Data](https://www.datarepository.movebank.org/handle/10255/move.882) via [Data is Plural](https://themarkup.org/data-is-plural/2023/01/25/from-jazz-solos-to-cats-on-the-move#:~:text=Giuseppe%20Sollazzo%5D-,Cats%20on%20the%20move,-.%20Between%202013).

Citation: Kays R, Dunn RR, Parsons AW, Mcdonald B, Perkins T, Powers S, Shell L, McDonald JL, Cole H, Kikillus H, Woods L, Tindle H, Roetman P (2020) The small home ranges and large local ecological impacts of pet cats. Animal Conservation. doi:10.1111/acv.12563

Citation: McDonald JL, Cole H (2020) Data from: The small home ranges and large local ecological impacts of pet cats \[United Kingdom\]. Movebank Data Repository. doi:10.5441/001/1.pf315732

```{r load libaries}
library(tidyverse)
library(geosphere)
library(ggtext)
library(showtext)
showtext_opts(dpi = 300)
showtext_auto(enable = TRUE)
```

```{r load font}
font_add_google("Archivo")
f1 = "Archivo"

font_add_google("Barlow", bold.wt = 500)
f2 = "Barlow"
```

```{r import data}
cats_uk <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk.csv')
cats_uk_reference <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-31/cats_uk_reference.csv')
```

```{r wrangle}
df = cats_uk |>
  filter(visible==TRUE) |>
  mutate(date=ymd(as.Date(timestamp))) 

df1 = df |>
  group_by(tag_id,date) |>
  arrange(timestamp, .by_group = TRUE) |>
  mutate(day_id = row_number(),
         long = lag(location_long), 
         lat = lag(location_lat)) |>
  select(date,timestamp, location_long, location_lat, long, lat,day_id,) |>
  ungroup() |>
  filter(day_id!=1) |>
  mutate(dist = distGeo(cbind(long, lat), cbind(location_long, location_lat)))

selected = df |>
  group_by(tag_id) |>
  summarise(ndays=n_distinct(date)) |>
  arrange(ndays) |>
  filter(ndays==1) |>
  pull(tag_id)

df2 = df1 |>
  filter(! tag_id %in% selected) |>
  group_by(tag_id, date) |>
  summarise(dist_d = sum(dist)) |>
  ungroup() |>
  group_by(tag_id) |>
  mutate(mean = mean(dist_d, na.rm=TRUE),
         med = median(dist_d, na.rm=TRUE),
         min=min(dist_d, na.rm=TRUE),
         max=max(dist_d, na.rm=TRUE))

df3 = df2 |>
  select(tag_id, mean) |>
  distinct() |>
  left_join(cats_uk_reference, by="tag_id") |>
  filter(animal_reproductive_condition %in% c("Spayed","Neutered"),
         !is.na(age_years)) |>
  ungroup() |>
  mutate(grp=case_when(age_years==0~"Kitten\n(up to 1 year)",
                       between(age_years,1,6)~"Young adult\n(1 - 6 years)",
                       between(age_years,7,10)~"Mature adult\n(7 - 10 years)",
                       age_years>10 ~"Senior\n(>10 years)"),
         grp = factor(grp, levels=c("Kitten\n(up to 1 year)","Young adult\n(1 - 6 years)","Mature adult\n(7 - 10 years)","Senior\n(>10 years)")),
         animal_reproductive_condition=case_when(animal_reproductive_condition=="Spayed"~"Spayed (female)", animal_reproductive_condition=="Neutered"~"Neutered (male)"),
         ncatlong= case_when(n_cats==1~"**1** cat",
                         n_cats>1~paste0("**",{n_cats},"**"," ","cats")))
```

```{r plot}
theme1= cowplot::theme_minimal_grid(12) +
  theme(text=element_text(family=f1),
        plot.margin=margin(.5,.7,.3,.5, unit="cm"),
        plot.caption.position = "plot",
        plot.caption = element_text(hjust=0, size=8.3, color="grey45", lineheight = 1.1, margin=margin(t=10)),
        plot.title.position = "plot",
        plot.title=element_text(size=16),
        plot.subtitle=element_text(lineheight=1.1, size=10,margin=margin(b=10), color="grey20"),
        axis.title = element_text(size=11),
        axis.text.y = element_markdown(hjust=0,margin=margin(l=3)),
        axis.text.x = element_text(margin=margin(b=3)),
        legend.title=element_text(size=10.8, face="bold"),
        legend.text=element_text( size=10.3),
        legend.position = "top",
        legend.box = 'vertical', 
        legend.box.just = 'left',
        legend.box.margin = margin(l=-45),
        legend.spacing.y=unit(.3,"lines"))
```

```{r, fig.height=3.5, fig.width=3.5}
df3 |>
  ggplot(aes(mean,fct_rev(ncatlong),
             shape=animal_reproductive_condition, fill=grp, order=grp)) +
  ggdist::geom_dots(group=NA, layout="hex", color="grey50") +
  scale_shape_manual(values=c(22,24)) +
  scale_fill_manual(values=c("#dd5129","#fab255","#43b284","#0f7ba2")) + #from MetBrewer
  theme1 +
  labs(x="Average distance travelled per day (meters)", 
       y="Number of cats in the house",
       fill="Age group:", color="Age group:", shape="Reproductive condition:",
       title="Pet cats on the move (UK)",
       subtitle="Estimate average (minimum) distance travelled per day by 93 neutered/spayed pet cats surveyed in the UK\nbetween June 2017 and November 2017. The distance in meters is an estimate of the shortest distance\nbetween two sensor measurement locations.",
       caption="#TidyTuesday week 5 | Data: Movebank for Animal Tracking Data via Data is Plural\nSources:Kays R, Dunn RR, Parsons AW, Mcdonald B, Perkins T, Powers S, Shell L, McDonald JL, Cole H, Kikillus H, Woods L,\nTindle H, Roetman P (2020) The small home ranges and large local ecological impacts of pet cats. Animal Conservation.\ndoi:10.1111/acv.12563. McDonald JL, Cole H (2020) Data from: The small home ranges and large local ecological impacts of\npet cats [United Kingdom]. Movebank Data Repository. doi:10.5441/001/1.pf315732") 

ggsave("export/2023_05.png", height=7, width=7, bg="#fafafa")
```
