---
title: "Tidy Tuesday Week 35/2021"
date: "2021/08/24"
output: html_notebook
---

[Tidy Tuesday](https://github.com/rfordatascience/tidytuesday) week 35
[Lemurs](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-08-24/readme.md), data from [Duke Lemur Center](https://lemur.duke.edu/) and was cleaned by [Jesse Mostipak](https://github.com/rfordatascience/tidytuesday/issues/204).

```{r}
# Load libraries
library(tidytuesdayR)
library(tidyverse)
library(scales)
library(ggtext)
library(gghalves)
library(DT)
library(rcartocolor)
library(lubridate)

options(dplyr.summarise.inform = FALSE)
theme_set(theme_minimal(base_size = 10))
```

```{r}
# import data
lemurs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv')
tax <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/taxonomy.csv')
```
```{r}
head(lemurs)
```
```{r}
# proportion missing data by variable
lemurs %>% summarise(across(everything(), ~mean(is.na(.)))) %>% 
  gather() %>%
  mutate(value=round(value,3)) %>%
  arrange(desc(value))

# unique values by variable
lemurs %>% summarise_all(n_distinct)
```

```{r}
# summary of record count per animal
lemurs %>% 
  count(dlc_id) %>% 
  summary()
```

### Weight by age category and gender group    
* shared on [Twitter](https://twitter.com/leeolney3/status/1430083827540578308)

```{r}
# Latest recorded weight by specimen ID
weight = lemurs %>% 
  mutate(age_cat = case_when(age_category=="young_adult" ~ "Young-Adult",
                             age_category=="IJ" ~ "Infant/Juvenile",
                             age_category=="adult" ~ "Adult"),
         gender= case_when(sex=="F"~ "Female",
                           sex=="M" ~ "Male")
         ) %>%
  mutate(group=paste(age_cat,gender)) %>%
  mutate(group=factor(group,levels=c("Adult Female","Adult Male",
                      "Young-Adult Female","Young-Adult Male",
                      "Infant/Juvenile Female","Infant/Juvenile Male","Infant/Juvenile NA"),
                      ordered=T))  %>% 
  group_by(dlc_id) %>%
  slice(which.max(weight_date)) 

# median
weight %>%
  group_by(group) %>%
  summarise(med=median(weight_g)) %>%
  arrange(desc(med))

# boxplot with barcode strips 
weight %>%
	ggplot(aes(y=weight_g, x=group, color=group)) + 
   geom_boxplot(
    width = .2, 
    outlier.shape = NA,
    fill="#f4f4f2",
    position=position_nudge(x=0.2,y=0)
  ) + 
  geom_point(
    shape = 95,
    size = 6,
    alpha = .2,
    position=position_nudge(x=-0.1,y=0)
  ) + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
  scale_color_manual(values=c("#003049","#003049",
                              "#f77f00","#003049",
                              "#003049","#003049","#003049")) + 
  theme(legend.position="none",
        panel.grid=element_line(size=.2, color="#ced4da"),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank(),
        axis.title.x=element_markdown(size=8.2, margin=margin(t=8)),
        axis.title.y=element_markdown(size=8.2, margin=margin(r=8)),
        axis.text.x=element_text(size=7.5, margin=margin(t=-5)),
        plot.caption=element_text(size=7, color="#495057"),
        plot.title.position = "plot",
        plot.title=element_text(face="bold", size=13),
        plot.subtitle = element_text(size=8),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        plot.background = element_rect(fill="#f4f4f2", color=NA)
        ) + 
  labs(x="**Age and gender group**", y="**Latest weight recorded** (in grams)",
       caption="#TidyTuesday Week 35 | Data from Duke Lemur Center",
       title="Weight of lemurs",
       subtitle="by age category and gender (latest date by specimen ID), where Young Adult Females have the highest median weight at 2155 grams.")

```

### Adult male weight by month of weigh-in
```{r}
# adult male: summary of record count
lemurs %>% 
  filter(sex=="M") %>%
  filter(age_category=="adult") %>%
  count(dlc_id) %>% summary(n)

# adult male weight by month of weigh-in
pal = colorRampPalette(rcartocolor::carto_pal(n=12,name="Prism")[1:10])
 
lemurs %>% 
  filter(sex=="M") %>%
  filter(age_category=="adult") %>%
  group_by(dlc_id) %>%
  mutate(n=n()) %>%
  filter(n>40) %>% 
  ggplot(aes(x=factor(birth_month), y=weight_g, color=factor(birth_month))) + 
  geom_half_boxplot(outlier.size = -1) + 
  geom_half_point(shape=21, alpha=0.4,size=1) +
  scale_color_manual(values=pal(n_distinct(lemurs$month_of_weight))) + 
  scale_y_continuous(breaks=seq(0,5000,1000)) +
  theme(legend.position="none",
        panel.grid=element_line(size=.3),
        panel.grid.minor=element_blank(),
        panel.grid.major.x=element_blank(),
        plot.title.position = "plot",
        axis.title.x=element_markdown(size=8.5),
        axis.title.y=element_markdown(size=8.5),
        plot.title=element_text(face="bold", size=13),
        plot.subtitle = element_text(size=8),
        plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        axis.text.x=element_text(margin=margin(t=-7, b=5))) + 
  labs(y="**Weight** (in grams)", x="**Month** (of the year the animal was weighed)",
       title="Weight of adult male lemurs",
       subtitle="by month of the year the animal was weighed, for adult male lemurs that have more than 40 total records")
```

### Lemur count by species
```{r}
tax = tax %>% 
  mutate(common_name= ifelse(common_name=="hybrid",
                            paste0(common_name," ","(",latin_name,")"),common_name),
         taxon= ifelse(taxon=="CMEAD","CMED", taxon))
```

```{r}
# join
common_name = lemurs %>% left_join(tax, by="taxon") %>%
  group_by(dlc_id) %>%
  slice(which.max(weight_date)) %>%
  ungroup() %>%
  mutate(common_name_group = fct_lump(common_name,20)) %>%
  count(common_name_group,sort=T) %>%
  mutate(common_name_group = forcats::fct_rev(forcats::fct_inorder(common_name_group)),
    common_name_group = forcats::fct_relevel(common_name_group, "Other", after = 0),
    perc = paste0(sprintf("%4.1f", n / sum(n) * 100), "%"),
    perc = if_else(row_number() == 1, paste(perc, "of all lemurs species"), perc),
    place = if_else(row_number() == 1, 1, 0.2),
    perc = paste(" ", perc, " ")
         )

pal <- c(
  "#cdd2d7",
  rep("#adb5bd", length(common_name$common_name_group) - 4), 
  "#b7094c", "#0091ad", "#f9c74f"
)

common_name %>%
  ggplot(aes(x=n, y=common_name_group, fill=common_name_group)) +
  geom_col() + 
  geom_text(aes(label=perc, hjust = place), size=2.8, fontface = "bold") + 
  scale_fill_manual(values = pal, guide = "none") + 
  scale_x_continuous(expand = c(.01, .01)) +
  theme_void(base_size = 10) + 
  theme(axis.text.y=element_text(hjust=1, size=8),
        plot.margin = margin(15, 30, 15, 15),
        plot.title.position = "plot",
        plot.title=element_text(face="bold", size=13,margin=margin(b=10))) +
  labs(title ="Lemur count by species")
```


### Weight of female lemurs compared to mal
```{r}
female = lemurs %>% left_join(tax, by="taxon") %>%
  filter(preg_status!="P") %>%
  filter(age_category!="IJ") %>%
  drop_na(common_name, sex, dlc_id, weight_g) %>%
  group_by(common_name, sex, dlc_id) %>%
  summarise(weight=mean(weight_g)) %>%
  ungroup() %>%
  group_by(common_name, sex) %>%
  summarise(weight2=mean(weight)) %>%
  mutate(diff=weight2-lag(weight2),
         col=ifelse(diff<0,"Female>Male","Male>Female")) %>%
  filter(sex=="M") %>%
  mutate(diff=-(diff),
         pct_diff=round(diff/weight2,4)) %>%
  ungroup()%>%
  mutate(al= ifelse(pct_diff==max(pct_diff)|pct_diff==min(pct_diff), 1,0.7))


ggplot(female,aes(y=reorder(common_name, pct_diff), x=pct_diff, fill=col)) + 
  geom_col(aes(alpha=al),show.legend=F) +
  geom_text(data=female %>% filter(col=="Female>Male"), aes(x=-0.003, y=reorder(common_name, pct_diff), label=common_name,alpha=al), hjust=1, size=2.6, color="#023e8a") +
  geom_text(data=female %>% filter(col=="Male>Female"), aes(x=0.003, y=reorder(common_name, pct_diff), label=common_name, alpha=al), hjust=0, size=2.6, color="#e85d04") +
  geom_vline(xintercept = 0, size=0.7) +
  scale_fill_manual(values=c("#023e8a","#e85d04")) +
  scale_x_continuous(limits=c(-.15,.15), breaks=seq(-.15,.15,.05), 
                     labels=scales::percent_format(accuracy = 1)) +
  scale_alpha_identity() +
  theme(panel.grid.minor=element_blank(),
        panel.grid.major.y=element_blank(),
        panel.grid=element_line(size=.3),
        axis.text.y=element_blank(),
        axis.text.x=element_text(size=8),
        axis.title=element_blank(),
        plot.margin = margin(15, 25, 15, 25),
        plot.title=element_text(size=10, face="bold"),
        plot.subtitle = element_text(size=8.7, margin=margin(b=10))
        ) + 
  labs(title="Weight of female lemurs compared to male",
       subtitle="Average non-pregnant female (adult and young adult) lemurs' weight expressed as a percentage difference to male")
```

### Age category, hybrid, birth type

```{r}
lemurs2 = lemurs %>% 
  mutate(age_category = case_when(age_category=="young_adult" ~ "Young-Adult",
                             age_category=="IJ" ~ "Infant/Juvenile",
                             age_category=="adult" ~ "Adult"),
         hybrid= case_when(hybrid=="N"~"Not hybrid",
                           hybrid=="Sp"~"Hybrid"),
         birth_type= case_when(birth_type=="CB"~'Captive-born',
                               birth_type=="WB"~'Wild-born',
                               birth_type=="Unk"~'Unknown'),
         gender= case_when(sex=="F"~ "Female",
                           sex=="M" ~ "Male")
         )

# age category
lemurs2 %>% 
  group_by(age_category) %>%
  summarise(n=n_distinct(dlc_id)) %>%
  arrange(desc(n)) 

# hybrid 
lemurs2 %>%
  group_by(hybrid) %>%
  summarise(n=n_distinct(dlc_id)) %>%
  arrange(desc(n))

# birth type
lemurs2 %>% 
  group_by(birth_type) %>%
  summarise(n=n_distinct(dlc_id)) %>%
  arrange(desc(n))

# birth type, age category of animals that are dead 
lemurs2 %>%
  filter(!is.na(dod)) %>%
  group_by(birth_type, age_category) %>%
  summarise(n=n_distinct(dlc_id)) %>%
  mutate(percentage=scales::percent(n / sum(n), accuracy = .1, trim = FALSE))
```

### Average weight of male captive-born lemurs
```{r}
# average weight of male captive born lemurs, by species and age category
lemurs %>% left_join(taxon_df, by="taxon") %>%
  filter(birth_type=="CB") %>%
  filter(sex=="M") %>%
  group_by(dlc_id, common_name, age_category) %>%
  summarise(weight=mean(weight_g)) %>%
  ungroup() %>%
  mutate(age_category = case_when(age_category=="young_adult" ~ "Young-Adult",
                             age_category=="IJ" ~ "Infant/Juvenile",
                             age_category=="adult" ~ "Adult")) %>%
  group_by(common_name, age_category) %>%
  summarise("average weight"= round(mean(weight),2)) %>%
  pivot_wider(names_from=age_category, values_from="average weight") %>%
  rename("Species"=common_name) %>%
  ungroup() %>%
  select("Species","Adult", "Young-Adult", "Infant/Juvenile") %>%
  DT::datatable(rownames=FALSE,options = list(order = list(list(1, 'desc'))),
                caption = htmltools::tags$caption( style = 'caption-side: top; text-align: center; color:black; font-size:130% ;','Average weight of male captive-born lemurs (in grams)'))
```

### Age category count by year
* reference: https://twitter.com/alyssastweeting/status/1430363273401044999
```{r}
age_year = lemurs2 %>% 
  mutate(age_category=factor(age_category,
                             levels=c("Infant/Juvenile","Young-Adult","Adult"), ordered=T)) %>%
  group_by(dlc_id) %>%
  mutate(weight_date = min(ymd(weight_date))) %>%
  mutate(year_joined=year(weight_date)) %>%
  summarise(name, taxon, year_joined, age_category) %>%
  unique() %>%
  group_by(year_joined, age_category) %>%
  mutate(n=length(year_joined)) %>%
  unique() 
```

```{r}
ggplot(age_year, aes(x=year_joined, fill=age_category)) + 
  geom_bar(stat="count", alpha=0.9) + 
  scale_fill_manual(values=c("#6699cc","#ff8c42","#a23e48"), guide="none") +
  scale_x_continuous(expand=c(0,0)) +
  theme(plot.title.position = "plot",
        panel.grid=element_line(size=.3),
        plot.subtitle=element_markdown(size=9.5),
        plot.title=element_text(face="bold", size=13),
        axis.title = element_text(size=8.5, face="bold"),
        plot.margin = margin(15, 25, 15, 25),
        ) + 
  labs(subtitle="Count of age categories by first record year <span style = 'color:#6699cc'><b>Infant/Juvenile</b></span>, <span style = 'color:#ff8c42'><b>Young-Adult</b></span> and <span style = 'color:#a23e48'><b>Adult</b></span>",
       title="Lemur age categories by year",
       x="Count", y="Year joined")
```

### Life span by species 
* reference: https://github.com/BlakeRMills/TidyTuesday/blob/main/Lemurs%20(24%20Aug%202021)/TidyTuesday%2024%20Aug%202021%20Lemurs.R

```{r, fig.width=3.5, fig.height=3.3}
life = lemurs %>% left_join(tax, by="taxon") %>%
  mutate(common_name = str_to_title(common_name) %>% str_remove(" Lemur"),
         birth_type= case_when(birth_type=="CB"~'Captive-born',
                               birth_type=="WB"~'Wild-born',
                               birth_type=="Unk"~'Unknown')) %>%
  group_by(dlc_id) %>%
  slice(which.max(weight_date)) %>%
  ungroup() %>%
  filter(!is.na(age_at_death_y)) %>%
  filter(birth_type!="Unknown")

life %>%
  group_by(common_name) %>%
  summarise(mean=mean(age_at_death_y)) %>%
  left_join(., life, by="common_name") %>%
  mutate(mean = round(mean, 1),
         label = sprintf('%.1f', mean)) %>%
  ggplot(aes(y=fct_rev(fct_reorder(common_name, desc(mean))), x=mean)) + 
  geom_point(aes(x= age_at_death_y, color=birth_type), shape=21) +
  geom_point(size=6, color="#5c4d7d") +
  geom_text(aes(label=mean),size=2, color="white") +
  scale_color_manual(values=c("#dda15e","#0077b6"), guide="none") +
  #coord_cartesian(expand=F, clip="off") +
  theme(panel.grid=element_line(size=.3),
        panel.grid.major.y=element_blank(),
        panel.grid.minor=element_blank(),
        plot.title.position = "plot",
        plot.margin = margin(15, 25, 15, 25),
        axis.title=element_text(size=8.5),
        axis.text.y=element_text(size=7.5,margin=margin(r=-10)),
        plot.title=element_text(size=10, face="bold"),
        plot.subtitle=element_markdown(size=8.5)
        ) +
  labs(x="Lifespan (in years)", y="Lemur species",
       title="Lemur lifespan by taxonomy",
       subtitle="<span style = 'color:#5c4d7d'><b>Average</b></span> lifespan of <span style = 'color:#dda15e'><b>Captive-born</b></span> and <span style = 'color:#0077b6'><b>Wild-born</b></span> lemurs by taxonomy")
```

### Number of offspring by year 
* reference: https://twitter.com/tacheboutit/status/1430257413307916293/photo/1

```{r, warning=F}
offspring = lemurs %>% 
  drop_na(dob,n_known_offspring) %>%
  mutate(year=year(dob)) %>%
  filter(between(year, 1958,2015)) %>%
  group_by(dlc_id) %>%
  slice(which.max(weight_date)) %>%
  ungroup() %>%
  group_by(year) %>% arrange(year) %>%
  mutate(sh = ifelse(n_known_offspring==max(n_known_offspring)|n_known_offspring==min(n_known_offspring)
                       ,19,1),
         al = ifelse(n_known_offspring==max(n_known_offspring)|n_known_offspring==min(n_known_offspring)
                       ,1,0.5)) %>%
  mutate(col= case_when(n_known_offspring==min(n_known_offspring)~"#4DBBD5FF",
                        n_known_offspring==max(n_known_offspring)~"#00A087FF",
                        TRUE~"grey50"))

summary = offspring %>% group_by(year) %>% summarise(med=median(n_known_offspring))
spline_df <- as.data.frame(spline(summary$year, summary$med))
```

```{r}
offspring %>%
  ggplot(aes(x=year, y= n_known_offspring)) +
  geom_rect(aes(xmin=1960, xmax=1970, ymin=0, ymax=Inf), fill="#f8f9fa", alpha=0.1, inherit.aes = F) +
  geom_rect(aes(xmin=1980, xmax=1990, ymin=0, ymax=Inf), fill="#f8f9fa", alpha=0.1, inherit.aes = F) +
  geom_rect(aes(xmin=2000, xmax=2010, ymin=0, ymax=Inf), fill="#f8f9fa", alpha=0.1, inherit.aes = F) +
  geom_point(aes(shape=sh, alpha=al, color=col)) +
  #geom_segment(data=summary, aes(x=year-0.5,xend=year+0.5, y=med, yend=med),size=1.5, color="red")
  #geom_step(data=summary, aes(x=year-0.5, y=med), color="red")
  geom_point(data=summary, aes(x=year, y=med), color="#E64B35FF") +
  geom_line(data=spline_df, aes(x=x,y=y), color="#E64B35FF") +
  scale_shape_identity() +
  scale_alpha_identity() +
  scale_color_identity() + 
  scale_x_continuous(breaks=seq(1960,2020,10), expand = c(.01, .01)) + 
  theme(panel.grid.minor.x=element_blank(),
        plot.title.position = "plot",
        axis.title = element_text(size=8.5, face="bold"),
        plot.title=element_text(size=12, face="bold"),
        plot.subtitle=element_markdown(size=8.5, margin=margin(b=10)),
        plot.margin = margin(15, 25, 15, 25),
        axis.text.x=element_text(margin=margin(t=-7))
        ) + 
  labs(x="Year",y="Number of offspring",
       title="Lemur offspring by year",
       subtitle="<span style = 'color:#E64B35FF'><b>Median</b></span>, <span style = 'color:#4DBBD5FF'><b>minimum</b></span> and <span style = 'color:#00A087FF'><b>maximum</b></span> count of offspring of individual is known to have produced (latest record date by specimen ID)")
```







